---
title: "Effect of biannual azithromycin distribution on antibody responses to malaria, bacterial, and protozoan pathogens in Niger"
subtitle: "Format GPS and climate data for analysis"
author: "Ben Arnold ben.arnold@ucsf.edu"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: haddock
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Preamble

```{r preamble}
library(here)
here()
#----------------------------
# source the config file
#----------------------------
source(here("R","mordor-ab-Config.R"))

#----------------------------
# spatial packages
# used only for this script
# (it's a pain to load them
# for all scripts b/c raster
# conflicts with dplyr::select)
#----------------------------
library(sp)
library(rgdal)
library(raster)
library(ncdf4)
```

# Load cluster list from morbidity data

Read in the final morbidity dataset shared by Jermey Keenan in December 2019. 

```{r format morbidity}
#-------------------------------
# read in the full morbidity file
#-------------------------------
md <- read_csv("~/Library/CloudStorage/Box-Box/mordor/mordor data/data/final-morbidity/MORDOR-Morbidity-merged-20Mar2019.csv",
                      col_types = cols(code_bspot_ = col_character(),
                                       arm = col_character(),
                                       txlabelgwu = col_character(),
                                       txdate = col_character(),
                                       dosex = col_character()))

#-------------------------------
# merge on the public IDs
# and drop study IDs to protect
# PII for public-facing data
#-------------------------------
public_ids <- read_csv("~/Library/CloudStorage/Box-Box/mordor/mordor data/data/final-morbidity/public-IDs/mordor-morbidity-public-IDs-child.csv")

md2 <- md %>%
  left_join(public_ids, by = c("gwu","txlabelgwu","masterperson_mcd"))
```

# Load MORDOR GPS

Load GPS centroids for MORDOR morbidity clusters

```{r load mordor gps}
#-------------------------------
# Load MORDOR morbidity cluster
# GPS coordinates
#-------------------------------
d_gps <- read_csv("~/Library/CloudStorage/Box-Box/mordor/mordor data/data/final-gps/mordor-morbidity-cluster-gps.csv")

#-------------------------------
# merge the cluster IDs
# onto the GPS coordinates
#-------------------------------
d_gps2 <- d_gps %>%
  left_join(md2, by = "gwu") %>%
  dplyr::select(clusterid_public,gwu,longitude,latitude)

```


# Load Climate Data

Monthly precipitation data, global extent

CRU TS4.04: Climatic Research Unit (CRU) Time-Series (TS) version 4.04 of high-resolution gridded data of month-by-month variation in climate (Jan. 1901- Dec. 2019)

Harris, I., Osborn, T. J., Jones, P. & Lister, D. Version 4 of the CRU TS monthly high-resolution gridded multivariate climate dataset. Sci Data 7, 109 (2020).
https://www.nature.com/articles/s41597-020-0453-3
https://www.ncbi.nlm.nih.gov/pubmed/32246091

http://data.ceda.ac.uk/badc/cru/data/cru_ts/cru_ts_4.04/data/pre
https://www.benjaminbell.co.uk/2018/01/getting-climate-data.html

```{r load CRU data}
#-------------------------------
# Load the CRU TS precipitation 
# dataset into R 
#-------------------------------
d_pre <- brick(here("data/cru_ts4_04_2011-2019","cru_ts4.04.2011.2019.pre.dat.nc"), varname="pre")
```

# Spatial merge of CRU to clusters

```{r spatial merge of precip data}
#-------------------------------
# Spatial merge of CRU time series 
# to the 30 MORDOR morbidity clusters
#-------------------------------
mordor_pre <- data.frame(clusterid_public=d_gps2$clusterid_public,
                         gwu=d_gps2$gwu,
                         extract(d_pre, d_gps2[c("longitude","latitude")])
) %>%
  filter(!is.na(clusterid_public))

#-------------------------------
# reshape the data from wide
# to long format
#-------------------------------
d_pre2 <- mordor_pre %>%
  group_by(clusterid_public,gwu) %>%
  pivot_longer(cols = starts_with("X"),names_to = "rawdate", names_prefix = "X", values_to = "pre")

#-------------------------------
# format the dates
# subset to one observation per cluster, per day
#-------------------------------
d_pre3 <- d_pre2 %>%
  mutate(pdate = as_date(rawdate))  %>%
  group_by(clusterid_public,pdate) %>%
  slice(1)

```

# Plot precipitation time series 

Plot the precipitation time series for the 30 communities

All communities share the same overall precipitation patterns.  Therefore, we can be confident in defining wet seasons consistently across communities.

```{r plot precip time series}
prets_plot <- ggplot(data = d_pre3, aes(x = pdate, y = pre, color = clusterid_public)) +
  geom_line(alpha = 0.3) +
  labs(y = "monthly precipitation (mm)", x = "date")

prets_plot
```

# Identify the wet season

Create indicators of "wet season", defined as >4 mm rainfall in each month. A 4 mm cutoff was chosen based on visual inspection of the classification across years in the 2011 - 2019 time series (below).

```{r code wet season}
d_pre4 <- d_pre3 %>%
  mutate(season = ifelse(pre>4, "wet season","dry season"),
         season = factor(season))
```

Plot of precipitation with seasons indicated

```{r plot precip time series coded as season}
prets_plot2 <- ggplot(data = d_pre4 %>% filter(clusterid_public=="community-1"), aes(x = pdate, y = pre)) +
  geom_polygon(alpha = 0.2) +
  geom_point(aes(color = season), alpha = 1, size = 0.7) +
  scale_color_manual(values = cbpal[c(2,3)],guide = guide_legend(override.aes = list(alpha=1, size = 2))) +
  scale_y_continuous(breaks = seq(0,250,by=50)) +
  scale_x_date() +
  labs(y = "monthly precipitation (mm)", x = "date") +
  theme_minimal()

prets_plot2

ggsave(here("figures","mordor-ab-monthly-precip.png"),device = "png", width = 8, height = 4)

```

# Save a precipitation dataset

Include monthly total precipitation estiamtes for each cluster, along with the estimated start and end date of the wet season in each year.

```{r save precipitation data}
#------------------------------------
# estimate the start and end of the
# wet season in each year
#------------------------------------
d_pre5 <- d_pre4 %>%
  ungroup() %>%
  mutate(pmnth = month(pdate),
         pyear = year(pdate)) %>%
  group_by(clusterid_public,pyear,season) %>%
  mutate(s_start = min(pdate),
         s_end = max(pdate),
         wet_start = ifelse(season=="wet season",s_start,NA),
         wet_end = ifelse(season == "wet season",s_end,NA)) %>%
  group_by(clusterid_public,pyear) %>%
  mutate(wet_start = as_date(max(wet_start, na.rm = TRUE)),
         wet_end = as_date(max(wet_end, na.rm = TRUE))
         ) %>%
  # restrict to relevant variables
  dplyr::select(clusterid_public,pdate,pyear,pmnth,season,wet_start,wet_end,pre) %>%
  arrange(clusterid_public,pdate) 

write_csv(d_pre5, file = here("data","mordor-ab-monthly-precip.csv"))
```

```{r detach spatial packages}
#------------------------------------
# detach spatial packages to avoid
# conflicts with dplyr::select
#------------------------------------
detach("package:ncdf4", unload=TRUE)
detach("package:raster", unload=TRUE)
detach("package:rgdal", unload=TRUE)
detach("package:sp", unload=TRUE)
```


# Session Info
```{r session info}
sessionInfo()
```


