---
title: "Effect of biannual azithromycin distribution on antibody responses to malaria, bacterial, and protozoan pathogens in Niger"
subtitle: "Assess baseline balance"
author: "Ben Arnold ben.arnold@ucsf.edu"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: haddock
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Summary

Estimate the number of children measured in each round to summarize in a CONSORT flow diagram.

# Preamble

```{r preamble}
library(here)
here()
#----------------------------
# source the config file
#----------------------------
source(here("R","mordor-ab-Config.R"))

#----------------------------
# source the functions file
# includes reused functions
#----------------------------
source(here("R","mordor-ab-Functions.R"))
```

# Load the antibody data

```{r load antibody data}
#-------------------------------
# load the formatted analysis
# data created by
# mordor-ab-format-data.Rmd
#-------------------------------
d <- read_rds(here("data","mordor-ab-analysis-public.rds"))

#-------------------------------
# drop "_public" suffix from IDs
#-------------------------------
d2 <- d %>%
  rename(clusterid = clusterid_public, childid = childid_public)


#-------------------------------
# remove CTB because it is so
# highly correlated with ETEC
# remove GST negative control
# remove trachoma and strongy
# because no transmission in this
# population
#-------------------------------
d2 <- d2 %>%
  filter(!antigen %in% c("ctb","gst","pgp3","ct694","nie")) %>%
  mutate(antigenf = factor(antigenf))
```

# Create composite seropositivity indicators

For enteric pathogens with multiple antigens (Cryptosporidium, Giardia, Salmonella, Campylobacter), create a series of seropositivity indicators that indicate if a child was positive to either antigen.

Then bind those to the original data frame to create a seropositivity dataset for analyses.

```{r composite seropositivity indicators}
d_enteric_comps <- d2 %>%
  filter(pathogen %in% c("Cryptosporidium","Giardia","Salmonella","Campylobacter")) %>%
  group_by(pathogen,childid,phase) %>%
  mutate(seropos = max(seropos)) %>%
  slice(1) %>%
  mutate(antigenf = case_when(
    antigen %in% c("cp17","cp23") ~ "Cryptosporidium Cp17 or Cp23",
    antigen %in% c("vsp3","vsp5") ~ "Giardia VSP-3 or VSP-5",
    antigen %in% c("salb","sald") ~ "Salmonella groups B or D LPS",
    antigen %in% c("p18","p39") ~ "Campylobacter p18 or p39"
  )) %>%
  select(clusterid,phase,arm,childid,agem,sex,dbsid,pathogen,antigenf,seropos)


```

```{r create seropositivity dataset, warning=FALSE}
dp <- d2 %>%
  # drop pathogens w/ composite indicators
  filter(!pathogen %in% c("Cryptosporidium","Giardia","Salmonella","Campylobacter")) %>%
  # add composite indicators
  bind_rows(d_enteric_comps) %>%
  # create new factor levels that includes composite indicators
  mutate(antigenf = factor(antigenf,levels = c(
    "Campylobacter p18 or p39",
    "ETEC LTB",
    "Salmonella groups B or D LPS",
    "Cryptosporidium Cp17 or Cp23",
    "Giardia VSP-3 or VSP-5",
   levels(d2$antigenf)[10:19]) )
         ) %>%
  # drop extra variables (MFI doesn't correspond to composite seropos so drop)
  select(-mfi,-logmfi)

```

# Reshape data for making table

Reshape the antibody data to wide format. Create age category variable

```{r reshape data wide}
dp_wide <- dp %>%
  select(clusterid,arm,phase,childid,agem,sex,antigenf,seropos) %>%
  mutate(seropos = ifelse(seropos == 1 , "Yes", "No"),
         seropos = factor(seropos)) %>%
  pivot_wider(id_cols = c("clusterid","arm","phase","childid","agem","sex"), names_from = antigenf, values_from = seropos) %>%
  mutate(agecat = cut(agem, breaks = c(-1,11,23,35,47,60)),
         agecat = factor(agecat, 
                         levels=c("(-1,11]","(11,23]","(23,35]","(35,47]","(47,60]"),
                         labels = c("<1 y", "1 y", "2 y", "3 y", "4 y")),
         arm = factor(arm, levels=c("placebo","azithro"),labels = c("Placebo","Azithromycin"))
         )

```


# Baseline balance table.

Note that for baseline seroprevalence estimates, the antibody responses are limited to the age ranges used in the primary analysis: malaria (12-59 months), enterics (6-24 months), Salmonella and Step (6-59 months)

```{r baseline balance table}

#----------------------------------
# Table for age and sex, 
# across all children at baseline
#----------------------------------
table1(~ agecat + sex  | arm, 
       data=dp_wide %>% filter(phase == 0), 
       overall = FALSE
       )

#----------------------------------
# Table for malaria seroprevalence
# ages 12-59 months
#----------------------------------
table1(~ `P. falciparum MSP-1` +
         `P. falciparum AMA1` +
         `P. falciparum GLURP-Ro` +
         `P. falciparum LSA1` + 
         `P. falciparum CSP` +
         `P. falciparum HRP2` +
         `P. malariae MSP-1` +
         `P. ovale MSP-1` +
         `P. vivax MSP-1` 
       | arm, 
       data=dp_wide %>% filter(phase == 0 & agem>=12), 
       overall = FALSE
       )

#----------------------------------
# Table for bacteria and protozoan 
# seroprevalence
# ages 6-24 months
#----------------------------------
table1(~ `ETEC LTB` | arm, 
       data=dp_wide %>% filter(phase == 0 & agem>=6 & agem<=24), 
       overall = FALSE
       )

#----------------------------------
# Table for salmonella and strep 
# seroprevalence
# ages 6-59 months
#----------------------------------
table1(~ `Campylobacter p18 or p39` +
         `Cryptosporidium Cp17 or Cp23` +
         `Giardia VSP-3 or VSP-5` +
         `Salmonella groups B or D LPS` +
         `Streptococcus group A SPEB` 
       | arm, 
       data=dp_wide %>% filter(phase == 0 & agem>=6), 
       overall = FALSE
       )


```

# Session Info
```{r session info}
sessionInfo()
```


