---
title: "Effect of biannual azithromycin distribution on antibody responses to malaria, bacterial, and protozoan pathogens in Niger"
subtitle: "Breadth of response to malaria antigens with composite analyses"
author: "Ben Arnold ben.arnold@ucsf.edu"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: haddock
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Summary

Assess the breadth of response to _P. falciparum_ antigens with composite analyses that combine information across the entire _P. falciparum_ panel.

# Preamble

```{r preamble}
library(here)
here()
#----------------------------
# source the config file
#----------------------------
source(here("R","mordor-ab-Config.R"))

#----------------------------
# source the functions file
# includes reused functions
#----------------------------
source(here("R","mordor-ab-Functions.R"))
```



# Load the antibody data

```{r load antibody data}
#-------------------------------
# load the formatted analysis
# data created by
# mordor-ab-format-data.Rmd
#-------------------------------
d <- read_rds(here("data","mordor-ab-analysis-public.rds"))

#-------------------------------
# drop "_public" suffix from IDs
#-------------------------------
d <- d %>%
  rename(clusterid = clusterid_public, childid = childid_public)

# filter to P. falciparum antigens
# filter to post-baseline measurements
d2 <- d %>%
  filter(antigen %in% c("pfmsp1","ama","glurp","lsa1","csp","hrp2")) %>%
  filter(phase > 0) %>%
  mutate(antigenf = factor(antigenf))


```

# Create composite seropositivity indicator

Create an indicator if a child is positive to any one of the P. falciparum antigens.  Also create a count of the number of antigens to which the child is positive.

```{r composite seropositivity indicator}
#-------------------------------
# create a composite seropositivity
# indicator across all P. falciparum
# antibody responses
#-------------------------------
d3 <- d2 %>%
  group_by(childid,phase) %>%
  mutate(nantigens = sum(seropos),
         seropos = max(seropos),
         ) %>%
  arrange(clusterid,childid,phase) %>%
  slice(1) %>%
  mutate(antigenf = "P. falciparum (MSP-1, AMA1, GLRUP-Ro, LSA1, CSP, HRP2)") %>%
  select(clusterid,phase,arm,childid,agem,sex,dbsid,pathogen,antigenf,seropos,nantigens)


```


# Estimate age-dependent seroprevalence by arm

```{r age-dependent seroprevalence by arm}

#-----------------------------
# create an indicator for 
# children living in intervention
# and control communities
#
# create a dummy indicator equal
# to 1 for all communities to zero
# out the marginal predictions
# over all community random effects
#
# create a factor variable
# for clusterid (for model)
#-----------------------------
d3 <- d3 %>%
  mutate(tr0 = ifelse(arm == "placebo",1,0),
         tr1 = ifelse(arm == "azithro",1,0),
         dummy = 1,
         clusteridf = factor(clusterid),
         phasef = factor(phase, levels=c("6","12","24","36"))
         )

#----------------------------------------
# fit a GAM model fitting separate
# splines by age for each arm
# estimate simultaneous 95% confidence 
# intervals
#----------------------------------------
fitp_age <- mgcv::gam(seropos ~ s(agem,bs="cr",by=tr1) + 
                        s(agem,bs="cr",by=tr0) + 
                         s(clusteridf,bs="re",by=dummy), 
                       family = binomial(link="cloglog"),
                       data=d3)
  newd0 <- d3 %>%  mutate(dummy=0, tr0 = 1, tr1 = 0)
  newd1 <- d3 %>%  mutate(dummy=0, tr1 = 1, tr0 = 0)
  fit_tr0_ci <- gamCI(m=fitp_age,newdata=newd0,nreps=1000)
  fit_tr1_ci <- gamCI(m=fitp_age,newdata=newd1,nreps=1000)
  fit_ci <- bind_rows(fit_tr0_ci,fit_tr1_ci)

# convert linear predictor to prevalence
fitp_age2 <- fit_ci %>%
  mutate(arm = ifelse(tr1==1,"azithro","placebo"),
         arm = factor(arm,levels = c("placebo","azithro"))) %>%
  mutate(fit = cloglogfn(fit),
         uprP = cloglogfn(uprP),
         lwrP = cloglogfn(lwrP),
         uprS = cloglogfn(uprS),
         lwrS = cloglogfn(lwrS),
         )  


```


# Age-seroprevalence figure

Seroprevalence figure 

```{r age-dependent seroprevalence by arm figure malaria, warning = FALSE}

#-----------------------------
# plot age-dependent seroprev
# by arm for malaria
#-----------------------------
fitp_malaria<- fitp_age2 %>%
  mutate(antigenf = factor(antigenf))

#-----------------------------
# create a bounding box
# to shade ages used in
# the primary analysis
# and in age-dependent FOI
# estimation
#-----------------------------
d_shade_box <- data.frame(antigenf = levels(fitp_malaria$antigenf)) %>%
  mutate(minX = 12, 
         maxX = 60,
         arm = "placebo",
         antigenf = factor(antigenf,levels(fitp_malaria$antigenf))
         )

pcols <- cbpal[c(2,6)]

plot_age_seroprev_malaria <- ggplot(data = fitp_malaria , aes(x = agem, y = fit, group = arm, color = arm, fill = arm)) +
  # add shade box
  geom_rect(data=d_shade_box, inherit.aes = FALSE, aes(xmin = minX, xmax = maxX, ymin = 0, ymax = 1), alpha=0.1) +
  # approximate pointwise 95% CIs
  # geom_ribbon(aes(ymin = lwrP, ymax = uprP), color = NA, alpha = 0.3) +
  # spline fits
  geom_line() +
  # plot aesthetics
  annotate("text",x=36,y=0.95,label="placebo",color=pcols[1],hjust=1) +
  annotate("text",x=36,y=0.75,label="azithromycin",color=pcols[2],hjust=0) +
  scale_y_continuous(breaks = seq(0,1,by=0.2), labels = sprintf("%1.0f",seq(0,1,by=0.2)*100)) +
  scale_x_continuous(breaks = seq(0,60,by=12), minor_breaks = seq(0,60,by=3)) +
  scale_color_manual(values = pcols) +
  scale_fill_manual(values = pcols) +
  coord_cartesian(ylim = c(0,1.04), xlim = c(0,60)) +
  labs(x = "Age, months", y = expression(paste(italic("P. falciparum")," IgG seroprevalence (%)"))) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
plot_age_seroprev_malaria

ggsave(here("figures","mordor-ab-age-curves-malaria-composite.pdf"),plot_age_seroprev_malaria, device = "pdf", width = 4, height = 4)

```


# Average effects of intervention

## Seroprevalence
Estimate average difference in seroprevalence and the relative reduction in force of infection. Limit the analysis to ages 12-59 months to avoid maternal IgG contributions.


```{r seroprev by arm}
#-----------------------------
# limit to children older than
# 12 months
#-----------------------------
d4 <- d3 %>%
  filter(agem >= 12)


#-----------------------------
# tally number of observations
# by antigen (including composites)
#-----------------------------
n_serop <- d4 %>%
  group_by(antigenf,arm) %>%
  mutate(nomiss = ifelse(!is.na(seropos),1,0)) %>%
  summarize(nobs = sum(nomiss), .groups = "keep") %>%
  pivot_wider(id_cols = c(antigenf), names_from = "arm", values_from = nobs) %>%
  rename(placebo_n = placebo,
         azithro_n = azithro)

#-----------------------------
# estimate seroprevalence by arm
# and difference in seroprevalence
#-----------------------------
mu_serop <- d4 %>%
  group_by(antigenf,arm) %>%
  summarize(mean_serop = mean(seropos), .groups = "keep") %>%
  pivot_wider(id_cols = c(antigenf), names_from = "arm", values_from = mean_serop) %>%
  mutate(diff = azithro - placebo)

#-----------------------------
# merge means onto the Ns
#-----------------------------
mu_serop <- left_join(n_serop, mu_serop, by=c("antigenf"))
```



```{r bootstrap mean seropos diff}
#-----------------------------
# bootstrap resample communities
# with replacement to estimate
# 95% CIs for difference in seroprev
#-----------------------------
set.seed(123)
nreps <- 1000
ncl <- length(unique(d4$clusterid))
bsamp <- matrix(sample(unique(d4$clusterid),size = ncl*nreps, replace = TRUE),nrow = ncl,ncol = nreps)
d_serop_boot <- d4 %>% ungroup() %>% select(clusterid,antigenf,arm,seropos)
mu_serop_boot <- foreach(bi = 1:nreps, .combine = rbind) %do% {
  
  if(bi %% 50 == 0) cat(bi,"\n") else cat(".")
  
  di <- left_join(data.frame(clusterid = bsamp[,bi], stringsAsFactors = FALSE),
                  d_serop_boot, by = "clusterid")
  di %>%
  group_by(antigenf,arm) %>%
  summarize(mean_serop = mean(seropos), .groups = "keep") %>%
  pivot_wider(id_cols = "antigenf", names_from = "arm", values_from = mean_serop) %>%
  mutate(diff = azithro - placebo, 
         bootrep = bi)

}

```

```{r mean seroprev bootstrap cis}
#-----------------------------
# calculate percentile 95%
# CIs from the bootstrap
# distribution merge to estimates
#-----------------------------
mu_serop_ci <- mu_serop_boot %>%
  group_by(antigenf) %>%
  mutate(mean_p_lb = quantile(placebo, probs = 0.025),
         mean_p_ub = quantile(placebo, probs = 0.975),
         mean_a_lb = quantile(azithro, probs = 0.025),
         mean_a_ub = quantile(azithro, probs = 0.975),
         diff_bmean = mean(diff),
         diff_bse = sd(diff),
         diff_lb = quantile(diff, probs = 0.025),
         diff_ub = quantile(diff, probs = 0.975)) %>%
  select(antigenf,starts_with("mean_"), starts_with("diff_")) %>%
  slice(1)
  
mu_serop2 <- mu_serop %>%
  left_join(mu_serop_ci, by = "antigenf")
  
```

```{r permutation p-values seroprev}
#-----------------------------
# estimate permutation p-values
# for the difference in seroprev
#
# pooling over phase 6+
#
# conduct the permutation
# test on unweighted cluster
# level means
#-----------------------------
clmeans_serop <- d4 %>% 
    group_by(clusterid,arm,antigenf) %>%
    summarize(mean_serop = mean(seropos), .groups = "keep")

set.seed(34527)
permute_serop <- foreach(abi = unique(d4$antigenf), .combine = rbind) %do% {
  di <- clmeans_serop %>% filter(antigenf == abi)
  pti <- oneway_test(mean_serop ~ arm, data= clmeans_serop,subset = clmeans_serop$antigenf==abi, distribution = "exact", alternative = "two.sided")
  data.frame(antigenf = abi, permute_p = pvalue(pti))
  }

```

```{r seroprev means table}
mu_serop_table <- mu_serop2 %>%
  ungroup() %>%
  left_join(permute_serop, by = "antigenf") %>%
  mutate(diff_ci = paste0(sprintf("%1.2f",diff)," (",sprintf("%1.2f",diff_lb),", ", sprintf("%1.2f",diff_ub),")"),
         permute_p = as.numeric(permute_p)) %>%
  select(antigenf,placebo, azithro, diff_ci, permute_p)

knitr::kable(mu_serop_table,
             col.names = c("Pathogen, antigen", "Placebo","Azithromycin","Difference (95% CI)","P-value*"), 
             digits = c(0,2,2,0,3),
             align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"),full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Proportion\nSeropositive" = 2,  " " = 2)) %>%
  footnote(
    symbol=c("Exact permutation P-value.")
  )

```

## Force of infection

Estimate serological force of infection through the seroconverion rate. Estimate the seroconversion rate from age-structured seroprevalence.

We use a generalized linear model with a complementary log-log link fit with current status, age-prevalence data, equivalent to a proportional hazards model. We model the baseline hazard in age using a semi-parametric, cubic spline, $g(\cdot)$, and include a binary indicator for treatment group. Let $A$ be a child's age and let $T$ be a binary indicator for azithromycin treatment. The model is:

\begin{equation}
  \log - \log(1-P(Y|A,T)) = g(A) + \beta X
\end{equation}


The hazard or force of infection at age $a$ is: $\lambda(a) = F'(a)/(1-F(a))$. From the definition of the hazard, we have: 

$1- F(a) = \exp [ - \int_0^a \lambda(a) da ]$ and $\int_0^a \lambda(a) da = - \log [1-F(a)]$.

This is the cumulative hazard. The average hazard per year (units of $a$) also divided $-\log[1-F(a)]$ by $a$.  So, all of the information is embedded in the cumulative distribution function, $F(a)$, which here is the age-dependent seroprevalence curve. 

The average hazard over the age period $a_1$ to $a_2$ is:

\begin{equation}
  \int_{a_1}^{a_2} \lambda(a) da = \frac{\log[1-F(a_1)]-\log[1-F(a_2)]}{a_2 - a_1}
\end{equation}

```{r FOI  estimates}


set.seed(123)

gfit <- gam(seropos ~ s(agem, bs="cr") + arm +
              s(clusteridf, bs = "re",by=dummy), 
            data=d4,
                family=binomial(link="cloglog"))
gsum <- summary(gfit)
    # estimate average seroconversion rate by arm
    # use a parametric bootstrap of the posterior VCOV matrix to get inference
    newd0 <- data.frame(agem=c(min(d4$agem),max(d4$agem)), arm = "placebo", dummy=0, clusteridf = "community-1")
    newd1 <- data.frame(agem=c(min(d4$agem),max(d4$agem)), arm = "azithro", dummy=0, clusteridf = "community-1")
    lambda0 <- avgFOI(gfit, newdata=newd0, a1=min(d4$agem),a2=max(d4$agem),nreps=10000)
    lambda1 <- avgFOI(gfit, newdata=newd1, a1=min(d4$agem),a2=max(d4$agem),nreps=10000)
    # estimate HR and 95% CI from the model from the model 
    hr <- as.numeric(exp(gfit$coefficients[2]))
    loghr_se <-as.numeric(gsum$se[2])
    # return the results
foi_ests <- data.frame(antigenf = unique(d4$antigenf),
                      lambda0 = lambda0$mufoi,
                      lambda0_se = lambda0$mufoi_se,
                      lambda0_min95 = lambda0$mufoi_lb,
                      lambda0_max95 = lambda0$mufoi_ub,
                      lambda1 = lambda1$mufoi,
                      lambda1_se = lambda1$mufoi_se,
                      lambda1_min95 = lambda1$mufoi_lb,
                      lambda1_max95 = lambda1$mufoi_ub,
                      hr = hr, 
                      loghr_se = loghr_se,
                      hr_min95 = exp(log(hr) - 1.96*loghr_se),
                      hr_max95 = exp(log(hr) + 1.96*loghr_se)
                      )


foi_ests_tab <- foi_ests %>%
  mutate(lambda0c = paste0(sprintf("%1.2f",lambda0*12)," (",
                            sprintf("%1.2f",lambda0_min95*12),", ", 
                            sprintf("%1.2f",lambda0_max95*12),")"),
         lambda1c = paste0(sprintf("%1.2f",lambda1*12)," (", 
                            sprintf("%1.2f",lambda1_min95*12),", ", 
                            sprintf("%1.2f",lambda1_max95*12),")"),
         hrc = paste0(sprintf("%1.2f",hr)," (",
                       sprintf("%1.2f",hr_min95),", ",
                       sprintf("%1.2f",hr_max95),")")
         )

knitr::kable(foi_ests_tab %>% select(antigenf,lambda0c,lambda1c,hrc),digits=3,
             caption="Average seroconversion rate (SCR) per year estimated from age-structured seroprevalence",
             align = "lccc",
             col.names = c("Antigens","Placebo SCR (95% CI)","Azithromycin SCR (95% CI)", "HR (95% CI)"),
             row.names = FALSE)  %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE)

```

# Analyses grouped by Ab durability

At the request of peer reviewers, create composite indicators separately for antigens thought to have more durable IgG responses (MSP-1, AMA1) and antigens thought to have less durable IgG responses (GLRUP-Ro, LSA1, CSP, HRP2).

## Create composite indicators

```{r composite indicators for short and long lived ab responses}
#-------------------------------
# create a composite seropositivity
# indicator across all long-lived
# P. falciparum antibody responses
#-------------------------------
d3long <- d2 %>%
  filter(antigen %in% c("pfmsp1","ama")) %>%
  group_by(childid,phase) %>%
  mutate(seropos = max(seropos)) %>%
  arrange(clusterid,childid,phase) %>%
  slice(1) %>%
  mutate(antigenf = "P. falciparum (MSP-1, AMA1)") %>%
  select(clusterid,phase,arm,childid,agem,sex,dbsid,pathogen,antigenf,seropos)

#-------------------------------
# create a composite seropositivity
# indicator across all short-lived
# P. falciparum antibody responses
#-------------------------------
d3short <- d2 %>%
  filter(antigen %in% c("glurp","lsa1","csp","hrp2")) %>%
  group_by(childid,phase) %>%
  mutate(seropos = max(seropos)) %>%
  arrange(clusterid,childid,phase) %>%
  slice(1) %>%
  mutate(antigenf = "P. falciparum (GLRUP-Ro, LSA1, CSP, HRP2)") %>%
  select(clusterid,phase,arm,childid,agem,sex,dbsid,pathogen,antigenf,seropos)


```

## Estimate age-dependent seroprevalence curves

```{r age-dependent seroprevalence by arm long}

#-----------------------------
# create an indicator for 
# children living in intervention
# and control communities
#
# create a dummy indicator equal
# to 1 for all communities to zero
# out the marginal predictions
# over all community random effects
#
# create a factor variable
# for clusterid (for model)
#-----------------------------
d3ls <- bind_rows(d3long,d3short) %>%
  mutate(tr0 = ifelse(arm == "placebo",1,0),
         tr1 = ifelse(arm == "azithro",1,0),
         dummy = 1,
         clusteridf = factor(clusterid),
         phasef = factor(phase, levels=c("6","12","24","36"))
         )
d3l <- d3ls %>% filter(antigenf=="P. falciparum (MSP-1, AMA1)")
d3s <- d3ls %>% filter(antigenf=="P. falciparum (GLRUP-Ro, LSA1, CSP, HRP2)")

#----------------------------------------
# fit a GAM model fitting separate
# splines by age for each arm
# estimate simultaneous 95% confidence 
# intervals, long-lived responses
#----------------------------------------
set.seed(24792)
fitp_age_long <- mgcv::gam(seropos ~ s(agem,bs="cr",by=tr1) + 
                        s(agem,bs="cr",by=tr0) + 
                         s(clusteridf,bs="re",by=dummy), 
                       family = binomial(link="cloglog"),
                       data=d3l)
  newd0 <- d3l %>%  mutate(dummy=0, tr0 = 1, tr1 = 0)
  newd1 <- d3l %>%  mutate(dummy=0, tr1 = 1, tr0 = 0)
  fit_tr0_ci <- gamCI(m=fitp_age_long,newdata=newd0,nreps=1000)
  fit_tr1_ci <- gamCI(m=fitp_age_long,newdata=newd1,nreps=1000)
  fit_ci <- bind_rows(fit_tr0_ci,fit_tr1_ci)

# convert linear predictor to prevalence
fitp_age_long2 <- fit_ci %>%
  mutate(arm = ifelse(tr1==1,"azithro","placebo"),
         arm = factor(arm,levels = c("placebo","azithro"))) %>%
  mutate(fit = cloglogfn(fit),
         uprP = cloglogfn(uprP),
         lwrP = cloglogfn(lwrP),
         uprS = cloglogfn(uprS),
         lwrS = cloglogfn(lwrS),
         )  

#----------------------------------------
# fit a GAM model fitting separate
# splines by age for each arm
# estimate simultaneous 95% confidence 
# intervals, shorter lived responses
#----------------------------------------
set.seed(543543)
fitp_age_short <- mgcv::gam(seropos ~ s(agem,bs="cr",by=tr1) + 
                        s(agem,bs="cr",by=tr0) + 
                         s(clusteridf,bs="re",by=dummy), 
                       family = binomial(link="cloglog"),
                       data=d3s)
  newd0 <- d3s %>%  mutate(dummy=0, tr0 = 1, tr1 = 0)
  newd1 <- d3s %>%  mutate(dummy=0, tr1 = 1, tr0 = 0)
  fit_tr0_ci <- gamCI(m=fitp_age_short,newdata=newd0,nreps=1000)
  fit_tr1_ci <- gamCI(m=fitp_age_short,newdata=newd1,nreps=1000)
  fit_ci <- bind_rows(fit_tr0_ci,fit_tr1_ci)

# convert linear predictor to prevalence
fitp_age_short2 <- fit_ci %>%
  mutate(arm = ifelse(tr1==1,"azithro","placebo"),
         arm = factor(arm,levels = c("placebo","azithro"))) %>%
  mutate(fit = cloglogfn(fit),
         uprP = cloglogfn(uprP),
         lwrP = cloglogfn(lwrP),
         uprS = cloglogfn(uprS),
         lwrS = cloglogfn(lwrS),
         )  

#----------------------------------------
# bind predictions together to make a
# single figure
#----------------------------------------
fitp_age_ls <- bind_rows(fitp_age_long2,fitp_age_short2)
```


## Age-seroprevalence figure

```{r age-dependent seroprevalence by arm figure malaria long and short, warning = FALSE}

#-----------------------------
# plot age-dependent seroprev
# by arm for malaria
#-----------------------------
fitp_malaria_ls <- fitp_age_ls %>%
  mutate(antigenf = factor(antigenf, levels=c("P. falciparum (MSP-1, AMA1)","P. falciparum (GLRUP-Ro, LSA1, CSP, HRP2)")))

#-----------------------------
# create a bounding box
# to shade ages used in
# the primary analysis
# and in age-dependent FOI
# estimation
#-----------------------------
d_shade_box <- data.frame(antigenf = levels(fitp_malaria_ls$antigenf)) %>%
  mutate(minX = 12, 
         maxX = 60,
         arm = "placebo",
         antigenf = factor(antigenf,levels(fitp_malaria_ls$antigenf))
         )

d_annotate_labs <- data.frame(x=c(36,36),y=c(0.95,0.75),hjust=c(1,0),
                              antigenf=factor("P. falciparum (MSP-1, AMA1)", 
                                              levels=levels(fitp_malaria_ls$antigenf)),
                              arm=c("placebo","azithro"),
                              labels=c("placebo","azithromycin"))

pcols <- cbpal[c(2,6)]

plot_age_seroprev_malaria_ls <- ggplot(data = fitp_malaria_ls , aes(x = agem, y = fit, group = arm, color = arm, fill = arm)) +
  facet_grid(cols=vars(antigenf)) +
  # add shade box
  geom_rect(data=d_shade_box, inherit.aes = FALSE, aes(xmin = minX, xmax = maxX, ymin = 0, ymax = 1), alpha=0.1) +
  # approximate pointwise 95% CIs
  # geom_ribbon(aes(ymin = lwrP, ymax = uprP), color = NA, alpha = 0.3) +
  # spline fits
  geom_line() +
  # plot aesthetics
  geom_text(data=d_annotate_labs, aes(x=x,y=y), label=d_annotate_labs$labels, hjust=d_annotate_labs$hjust) +
  # annotate("text",x=36,y=0.95,label="placebo",color=pcols[1],hjust=1) +
  # annotate("text",x=36,y=0.75,label="azithromycin",color=pcols[2],hjust=0) +
  scale_y_continuous(breaks = seq(0,1,by=0.2), labels = sprintf("%1.0f",seq(0,1,by=0.2)*100)) +
  scale_x_continuous(breaks = seq(0,60,by=12), minor_breaks = seq(0,60,by=3)) +
  scale_color_manual(values = pcols) +
  scale_fill_manual(values = pcols) +
  coord_cartesian(ylim = c(0,1.04), xlim = c(0,60)) +
  labs(x = "Age, months", y = expression(paste(italic("P. falciparum")," IgG seroprevalence (%)"))) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
plot_age_seroprev_malaria_ls

ggsave(here("figures","mordor-ab-age-curves-malaria-composite-long-short.pdf"),plot_age_seroprev_malaria_ls, device = "pdf", width = 8, height = 4)

```


## Average effects of intervention

### Seroprevalence
Estimate average difference in seroprevalence and the relative reduction in force of infection. Limit the analysis to ages 12-59 months to avoid maternal IgG contributions.


```{r seroprev by arm by antigen groups}
#-----------------------------
# limit to children older than
# 12 months
#-----------------------------
d4ls <- d3ls %>%
  filter(agem >= 12) %>%
  mutate(antigenf = factor(antigenf, levels=c("P. falciparum (MSP-1, AMA1)","P. falciparum (GLRUP-Ro, LSA1, CSP, HRP2)")))

#-----------------------------
# tally number of observations
# by antigen (including composites)
#-----------------------------
n_serop <- d4ls %>%
  group_by(antigenf,arm) %>%
  mutate(nomiss = ifelse(!is.na(seropos),1,0)) %>%
  summarize(nobs = sum(nomiss), .groups = "keep") %>%
  pivot_wider(id_cols = c(antigenf), names_from = "arm", values_from = nobs) %>%
  rename(placebo_n = placebo,
         azithro_n = azithro)

#-----------------------------
# estimate seroprevalence by arm
# and difference in seroprevalence
#-----------------------------
mu_serop <- d4ls %>%
  group_by(antigenf,arm) %>%
  summarize(mean_serop = mean(seropos), .groups = "keep") %>%
  pivot_wider(id_cols = c(antigenf), names_from = "arm", values_from = mean_serop) %>%
  mutate(diff = azithro - placebo)

#-----------------------------
# merge means onto the Ns
#-----------------------------
mu_serop <- left_join(n_serop, mu_serop, by=c("antigenf"))
```

```{r bootstrap mean seropos diff by antigen groups}
#-----------------------------
# bootstrap resample communities
# with replacement to estimate
# 95% CIs for difference in seroprev
#-----------------------------
set.seed(123)
nreps <- 1000
ncl <- length(unique(d4ls$clusterid))
bsamp <- matrix(sample(unique(d4ls$clusterid),size = ncl*nreps, replace = TRUE),nrow = ncl,ncol = nreps)
d_serop_boot <- d4ls %>% ungroup() %>% select(clusterid,antigenf,arm,seropos)
mu_serop_boot <- foreach(bi = 1:nreps, .combine = rbind) %do% {
  
  if(bi %% 50 == 0) cat(bi,"\n") else cat(".")
  
  di <- left_join(data.frame(clusterid = bsamp[,bi], stringsAsFactors = FALSE),
                  d_serop_boot, by = "clusterid")
  di %>%
  group_by(antigenf,arm) %>%
  summarize(mean_serop = mean(seropos), .groups = "keep") %>%
  pivot_wider(id_cols = "antigenf", names_from = "arm", values_from = mean_serop) %>%
  mutate(diff = azithro - placebo, 
         bootrep = bi)

}

```

```{r mean seroprev bootstrap cis by antigen groups}
#-----------------------------
# calculate percentile 95%
# CIs from the bootstrap
# distribution merge to estimates
#-----------------------------
mu_serop_ci <- mu_serop_boot %>%
  group_by(antigenf) %>%
  mutate(mean_p_lb = quantile(placebo, probs = 0.025),
         mean_p_ub = quantile(placebo, probs = 0.975),
         mean_a_lb = quantile(azithro, probs = 0.025),
         mean_a_ub = quantile(azithro, probs = 0.975),
         diff_bmean = mean(diff),
         diff_bse = sd(diff),
         diff_lb = quantile(diff, probs = 0.025),
         diff_ub = quantile(diff, probs = 0.975)) %>%
  select(antigenf,starts_with("mean_"), starts_with("diff_")) %>%
  slice(1)
  
mu_serop2 <- mu_serop %>%
  left_join(mu_serop_ci, by = "antigenf")
  
```

```{r permutation p-values seroprev by antigen groups}
#-----------------------------
# estimate permutation p-values
# for the difference in seroprev
#
# pooling over phase 6+
#
# conduct the permutation
# test on unweighted cluster
# level means
#-----------------------------
clmeans_serop <- d4ls %>% 
    group_by(clusterid,arm,antigenf) %>%
    summarize(mean_serop = mean(seropos), .groups = "keep")

set.seed(34527)
permute_serop <- foreach(abi = unique(d4ls$antigenf), .combine = rbind) %do% {
  di <- clmeans_serop %>% filter(antigenf == abi)
  pti <- oneway_test(mean_serop ~ arm, data= clmeans_serop,subset = clmeans_serop$antigenf==abi, distribution = "exact", alternative = "two.sided")
  data.frame(antigenf = abi, permute_p = pvalue(pti))
  }

```

```{r seroprev means table by antigen groups}
mu_serop_table_ls <- mu_serop2 %>%
  ungroup() %>%
  left_join(permute_serop, by = "antigenf") %>%
  mutate(diff_ci = paste0(sprintf("%1.2f",diff)," (",sprintf("%1.2f",diff_lb),", ", sprintf("%1.2f",diff_ub),")"),
         permute_p = as.numeric(permute_p),
         antigenf=factor(antigenf,levels=levels(d4ls$antigenf))) %>%
  select(antigenf,placebo, azithro, diff_ci, permute_p)

knitr::kable(mu_serop_table_ls,
             col.names = c("Pathogen, antigen", "Placebo","Azithromycin","Difference (95% CI)","P-value*"), 
             digits = c(0,2,2,0,3),
             align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"),full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Proportion\nSeropositive" = 2,  " " = 2)) %>%
  footnote(
    symbol=c("Exact permutation P-value.")
  )

```
### Force of infection
```{r FOI  estimates by antigen groups}
d4l <- d4ls %>% filter(antigenf=="P. falciparum (MSP-1, AMA1)")
d4s <- d4ls %>% filter(antigenf=="P. falciparum (GLRUP-Ro, LSA1, CSP, HRP2)")
#-----------------------------
# estimate FOI and relative
# hazard for longer-lived
# antibody responses
#-----------------------------
set.seed(123)
gfit_long <- gam(seropos ~ s(agem, bs="cr") + arm +
              s(clusteridf, bs = "re",by=dummy), 
            data=d4l,
                family=binomial(link="cloglog"))
gsum <- summary(gfit_long)
    # estimate average seroconversion rate by arm
    # use a parametric bootstrap of the posterior VCOV matrix to get inference
    newd0 <- data.frame(agem=c(min(d4l$agem),max(d4l$agem)), arm = "placebo", dummy=0, clusteridf = "community-1")
    newd1 <- data.frame(agem=c(min(d4l$agem),max(d4l$agem)), arm = "azithro", dummy=0, clusteridf = "community-1")
    lambda0 <- avgFOI(gfit_long, newdata=newd0, a1=min(d4l$agem),a2=max(d4l$agem),nreps=10000)
    lambda1 <- avgFOI(gfit_long, newdata=newd1, a1=min(d4l$agem),a2=max(d4l$agem),nreps=10000)
    # estimate HR and 95% CI from the model from the model 
    hr <- as.numeric(exp(gfit_long$coefficients[2]))
    loghr_se <-as.numeric(gsum$se[2])
    # return the results
foi_ests <- data.frame(antigenf = unique(d4l$antigenf),
                      lambda0 = lambda0$mufoi,
                      lambda0_se = lambda0$mufoi_se,
                      lambda0_min95 = lambda0$mufoi_lb,
                      lambda0_max95 = lambda0$mufoi_ub,
                      lambda1 = lambda1$mufoi,
                      lambda1_se = lambda1$mufoi_se,
                      lambda1_min95 = lambda1$mufoi_lb,
                      lambda1_max95 = lambda1$mufoi_ub,
                      hr = hr, 
                      loghr_se = loghr_se,
                      hr_min95 = exp(log(hr) - 1.96*loghr_se),
                      hr_max95 = exp(log(hr) + 1.96*loghr_se)
                      )


foi_ests_tabl <- foi_ests %>%
  mutate(lambda0c = paste0(sprintf("%1.2f",lambda0*12)," (",
                            sprintf("%1.2f",lambda0_min95*12),", ", 
                            sprintf("%1.2f",lambda0_max95*12),")"),
         lambda1c = paste0(sprintf("%1.2f",lambda1*12)," (", 
                            sprintf("%1.2f",lambda1_min95*12),", ", 
                            sprintf("%1.2f",lambda1_max95*12),")"),
         hrc = paste0(sprintf("%1.2f",hr)," (",
                       sprintf("%1.2f",hr_min95),", ",
                       sprintf("%1.2f",hr_max95),")")
         )


#-----------------------------
# estimate FOI and relative
# hazard for shorter-lived
# antibody responses
#-----------------------------
set.seed(3243)
gfit_short <- gam(seropos ~ s(agem, bs="cr") + arm +
              s(clusteridf, bs = "re",by=dummy), 
            data=d4s,
                family=binomial(link="cloglog"))
gsum <- summary(gfit_short)
    # estimate average seroconversion rate by arm
    # use a parametric bootstrap of the posterior VCOV matrix to get inference
    newd0 <- data.frame(agem=c(min(d4s$agem),max(d4s$agem)), arm = "placebo", dummy=0, clusteridf = "community-1")
    newd1 <- data.frame(agem=c(min(d4s$agem),max(d4s$agem)), arm = "azithro", dummy=0, clusteridf = "community-1")
    lambda0 <- avgFOI(gfit_short, newdata=newd0, a1=min(d4s$agem),a2=max(d4s$agem),nreps=10000)
    lambda1 <- avgFOI(gfit_short, newdata=newd1, a1=min(d4s$agem),a2=max(d4s$agem),nreps=10000)
    # estimate HR and 95% CI from the model from the model 
    hr <- as.numeric(exp(gfit_short$coefficients[2]))
    loghr_se <-as.numeric(gsum$se[2])
    # return the results
foi_ests <- data.frame(antigenf = unique(d4s$antigenf),
                      lambda0 = lambda0$mufoi,
                      lambda0_se = lambda0$mufoi_se,
                      lambda0_min95 = lambda0$mufoi_lb,
                      lambda0_max95 = lambda0$mufoi_ub,
                      lambda1 = lambda1$mufoi,
                      lambda1_se = lambda1$mufoi_se,
                      lambda1_min95 = lambda1$mufoi_lb,
                      lambda1_max95 = lambda1$mufoi_ub,
                      hr = hr, 
                      loghr_se = loghr_se,
                      hr_min95 = exp(log(hr) - 1.96*loghr_se),
                      hr_max95 = exp(log(hr) + 1.96*loghr_se)
                      )


foi_ests_tabs <- foi_ests %>%
  mutate(lambda0c = paste0(sprintf("%1.2f",lambda0*12)," (",
                            sprintf("%1.2f",lambda0_min95*12),", ", 
                            sprintf("%1.2f",lambda0_max95*12),")"),
         lambda1c = paste0(sprintf("%1.2f",lambda1*12)," (", 
                            sprintf("%1.2f",lambda1_min95*12),", ", 
                            sprintf("%1.2f",lambda1_max95*12),")"),
         hrc = paste0(sprintf("%1.2f",hr)," (",
                       sprintf("%1.2f",hr_min95),", ",
                       sprintf("%1.2f",hr_max95),")")
         )

foi_ests_tab_ls <- bind_rows(foi_ests_tabl,foi_ests_tabs)


knitr::kable(foi_ests_tab_ls %>% select(antigenf,lambda0c,lambda1c,hrc),digits=3,
             caption="Average seroconversion rate (SCR) per year estimated from age-structured seroprevalence",
             align = "lccc",
             col.names = c("Antigens","Placebo SCR (95% CI)","Azithromycin SCR (95% CI)", "HR (95% CI)"),
             row.names = FALSE)  %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE)

```



# Estimate the cumulative number of antigens positive by age

```{r plot n antigen positive by age 12-36 months}

d_agecat <- d3 %>%
  mutate(agecat = cut(agem,breaks = c(-0.1,12,24,36,48,60)),
         agecat = factor(agecat,levels = c("(-0.1,12]","(12,24]","(24,36]","(36,48]",  "(48,60]"), labels = c("<1", "1","2","3","4"))) 

pcols <- cbpal[c(2,6)]
plot_nantigens_by_age <- ggplot(data = d_agecat, aes(x = agecat, y = nantigens,  color = arm, fill = arm)) +
  # geom_point(alpha = 0.1, color = "black") +
  # geom_smooth(se = FALSE) +
  # geom_jitter(aes(group = arm),width=0.1,alpha=0.1) +
  geom_boxplot(width = 0.5, outlier.shape = NA,alpha = 0.3) +
  scale_y_continuous(breaks = 0:6) +
  scale_color_manual(values = pcols) +
  scale_fill_manual(values = pcols) +
  labs(x = "age (years completed)", y = "number of P. falciparum antigens positive") +
  coord_cartesian(ylim=c(0,6)) +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank()
  )

plot_nantigens_by_age



```

# Session Info
```{r session info}
sessionInfo()
```


