---
title: "Biannual azithromycin distribution and antibody responses to malaria, bacterial, and protozoan pathogens among Nigerien children"
subtitle: "Sensitivity analysis that varies cutoff values +/- 20%"
author: "Ben Arnold ben.arnold@ucsf.edu"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: haddock
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Summary

Repeat the primary analysis that compares azithromycin and placebo groups by seropositivity and seroconversion rates, varying the seropositivity cutoff used in the primary analysis by +/- 20%.

# Preamble

```{r preamble, message = FALSE}
library(here)
here()
#----------------------------
# source the config file
#----------------------------
source(here("R","mordor-ab-Config.R"))

#----------------------------
# source the functions file
# includes reused functions
#----------------------------
source(here("R","mordor-ab-Functions.R"))
```



# Load the antibody data

```{r load antibody data}
#-------------------------------
# load the formatted analysis
# data created by
# mordor-ab-format-data.Rmd
#-------------------------------
d <- read_rds(here("data","mordor-ab-analysis-public.rds"))

#-------------------------------
# drop "_public" suffix from IDs
#-------------------------------
d <- d %>%
  rename(clusterid = clusterid_public, childid = childid_public)

#-------------------------------
# drop GST negative control values
# drop trachoma and strongy because
# no transmission in this population
#-------------------------------
d1 <- d %>%
  filter(!antigen %in% c("gst","pgp3","ct694","nie")) %>%
  mutate(antigenf = factor(antigenf))

#-------------------------------
# drop cholera CTB because so 
# highly cross-reactive with ETEC LTB
#-------------------------------
d1 <- d1 %>%
  filter( !antigen %in% c("ctb") ) %>%
  mutate(antigenf = factor(antigenf))

#-------------------------------
# drop baseline measurements
#-------------------------------
d2 <- d1 %>%
  filter(phase > 0)

#-------------------------------
# age restrictions based on 
# age-antibody curves
# to exclude ages with evidence
# for maternal IgG contributions
#
# Bacteria and protozoa: 6 mo - 59 mo
#
# ETEC LTB: 6 mo - 24 mo (no heterogeneity at older ages)
#
# Malaria antigens: 12 - 59 mo
#-------------------------------
d2 <- d2 %>%
  filter(agem >= 6) %>%
  mutate(agedrop = case_when(
    antigen %in% c("etec") & (agem > 24) ~ 1,
    antigen %in% c("hrp2","csp","glurp","lsa1","ama","pfmsp1","pvmsp1","pomsp1","pmmsp1") & (agem < 12) ~ 1,
    TRUE ~ 0
  )
         ) %>%
  filter(agedrop == 0)

```

# Create sensitivity seropositivity indicators

Create seropositivity indicators based on increasing and decreasing the seropositivity cutoff values by 20%.

```{r sensitivity seropositivity cutoffs}
d3 <- d2 %>%
  mutate(serocut_plus20 = serocut*1.2,
         serocut_minus20 = serocut*0.8,
         seropos_plus20 = ifelse(logmfi > log10(serocut_plus20),1,0 ),
         seropos_minus20 = ifelse(logmfi > log10(serocut_minus20),1,0 ),)
```

# Create composite seropositivity indicators

For enteric pathogens with multiple antigens (Cryptosporidium, Giardia, Salmonella, Campylobacter), create a series of seropositivity indicators that indicate if a child was positive to either antigen.

Then bind those to the original data frame to create a seropositivity dataset for analyses.

```{r composite seropositivity indicators}
d_enteric_comps <- d3 %>%
  filter(pathogen %in% c("Cryptosporidium","Giardia","Salmonella","Campylobacter")) %>%
  group_by(pathogen,childid,phase) %>%
  mutate(seropos=max(seropos),
         seropos_plus20 = max(seropos_plus20),
         seropos_minus20 = max(seropos_minus20)) %>%
  slice(1) %>%
  mutate(antigenf = case_when(
    antigen %in% c("cp17","cp23") ~ "Cryptosporidium Cp17 or Cp23",
    antigen %in% c("vsp3","vsp5") ~ "Giardia VSP-3 or VSP-5",
    antigen %in% c("salb","sald") ~ "Salmonella groups B or D LPS",
    antigen %in% c("p18","p39") ~ "Campylobacter p18 or p39"
  )) %>%
  select(clusterid,phase,arm,childid,agem,sex,dbsid,pathogen,antigenf,seropos,seropos_plus20, seropos_minus20)


```

```{r create seropositivity dataset, warning=FALSE}
d_serop <- d3 %>%
  # drop pathogens w/ composite indicators
  filter(!pathogen %in% c("Cryptosporidium","Giardia","Salmonella","Campylobacter")) %>%
  # add composite indicators
  bind_rows(d_enteric_comps) %>%
  # create new factor levels that includes composite indicators
  mutate(antigenf = factor(antigenf,levels = c(
    "Campylobacter p18 or p39",
    "ETEC LTB",
    "Salmonella groups B or D LPS",
    "Cryptosporidium Cp17 or Cp23",
    "Giardia VSP-3 or VSP-5",
   levels(d2$antigenf)[10:19]) )
         ) %>%
  # drop extra variables (MFI doesn't correspond to composite seropos so drop)
  dplyr::select(-mfi,-logmfi,-agedrop)

```

# Summarize the number of measurements reclassified by pathogen

Summarize measurements by pathogen and seropositivity status depending on the cutoff used.

```{r n by pathogen and seropositivity cutoff}
#-----------------------------
# summarize number seropositive
# by different cutoffs
# and the difference in number
# of seropositive after varying
# the cutoff values
#-----------------------------
d_n <- d_serop %>%
  group_by(antigenf) %>%
  summarize(nobs = n(),
            seropos = sum(seropos),
            seropos_minus20 = sum(seropos_minus20),
            seropos_plus20 = sum(seropos_plus20), 
            .groups = "keep") %>%
  mutate(diff_minus20 = seropos_minus20 - seropos,
         diff_plus20 = seropos_plus20 - seropos,
         seroprev = seropos/nobs,
         seroprev_minus20 = seropos_minus20/nobs,
         seroprev_plus20 = seropos_plus20/nobs) %>%
  # create formatted columms for a table
  mutate(seroprevf = paste0(format(seropos,big.mark = ",")," (",sprintf("%1.1f",seroprev*100),"%)"),
         seroprev_minus20f = paste0(format(seropos_minus20,big.mark = ",")," (",sprintf("%1.1f",seroprev_minus20*100),"%)"),
         seroprev_plus20f = paste0(format(seropos_plus20,big.mark = ",")," (",sprintf("%1.1f",seroprev_plus20*100),"%)")
           )

```

```{r table of seroprev by pathogen adjusting the cutoff}
knitr::kable(d_n %>% select(antigenf, nobs, seroprevf, seroprev_minus20f, seroprev_plus20f), 
             col.names = c("Pathogen","N", "Seropositive (%)", "Seropositive (%)", "Seropositive (%)")) %>%
  add_header_above(c(" " = 2, "Primary Analysis" = 1, "MFI Cutoff +20%" = 1, "MFI Cutoff -20%" = 1)) %>%
  kable_styling(bootstrap_options = "striped")

```


# Seroprevalence

## Estimate means by arm
```{r seroprev by arm}

#-----------------------------
# tally number of observations
# by antigen (including composites)
#-----------------------------
n_serop <- d_serop %>%
  group_by(pathogen,antigenf,arm) %>%
  mutate(nomiss = ifelse(!is.na(seropos),1,0)) %>%
  summarize(nobs = sum(nomiss), .groups = "keep") %>%
  pivot_wider(id_cols = c(pathogen, antigenf), names_from = "arm", values_from = nobs) %>%
  rename(placebo_n = placebo,
         azithro_n = azithro)

#-----------------------------
# estimate seroprevalence by arm
# and difference in seroprevalence
#-----------------------------
mu_serop <- d_serop %>%
  group_by(pathogen,antigenf,arm) %>%
  summarize(mean_serop = mean(seropos), .groups = "keep") %>%
  pivot_wider(id_cols = c(pathogen, antigenf), names_from = "arm", values_from = mean_serop) %>%
  mutate(diff = azithro - placebo)

mu_serop_minus20 <- d_serop %>%
  group_by(pathogen,antigenf,arm) %>%
  summarize(mean_serop_minus20 = mean(seropos_minus20), .groups = "keep") %>%
  pivot_wider(id_cols = c(pathogen, antigenf), names_from = "arm", values_from = mean_serop_minus20) %>%
  mutate(diff_minus20 = azithro - placebo) %>%
  rename(azithro_minus20 = azithro, placebo_minus20 = placebo)

mu_serop_plus20 <- d_serop %>%
  group_by(pathogen,antigenf,arm) %>%
  summarize(mean_serop_plus20 = mean(seropos_plus20), .groups = "keep") %>%
  pivot_wider(id_cols = c(pathogen, antigenf), names_from = "arm", values_from = mean_serop_plus20) %>%
  mutate(diff_plus20 = azithro - placebo) %>%
  rename(azithro_plus20 = azithro, placebo_plus20 = placebo)

#-----------------------------
# merge means onto the Ns
#-----------------------------
mu_serop <- left_join(n_serop, mu_serop, by=c("pathogen","antigenf")) %>%
  left_join(mu_serop_minus20, by = c("pathogen","antigenf")) %>%
  left_join(mu_serop_plus20, by = c("pathogen","antigenf"))
            
```

## Bootstrap 95% CIs
```{r bootstrap mean seropos diff}
#-----------------------------
# bootstrap resample communities
# with replacement to estimate
# 95% CIs for difference in seroprev
#-----------------------------
nreps <- 1000
ncl <- length(unique(d_serop$clusterid))
bsamp <- matrix(sample(unique(d_serop$clusterid),size = ncl*nreps, replace = TRUE),nrow = ncl,ncol = nreps)
d_serop_boot <- d_serop %>% ungroup %>% select(clusterid,antigenf,arm,seropos,seropos_minus20,seropos_plus20)

mu_serop_boot <- foreach(bi = 1:nreps, .combine = rbind) %dopar% {
  set.seed(bi)
  # if(bi %% 50 == 0) cat(bi,"\n") else cat(".")
  
  di <- left_join(data.frame(clusterid = bsamp[,bi], stringsAsFactors = FALSE),
                  d_serop_boot, by = "clusterid")
  di %>%
  group_by(antigenf,arm) %>%
  summarize(seroprev = mean(seropos), 
            seroprev_minus20 = mean(seropos_minus20),
            seroprev_plus20 = mean(seropos_plus20),
            .groups = "keep") %>%
    mutate(bootrep = bi)
}
```

```{r mean seroprev bootstrap cis}
#-----------------------------
# calculate percentile 95%
# CIs from the bootstrap
# distribution merge to estimates
#-----------------------------
mu_serop_ci <- mu_serop_boot %>%
  group_by(antigenf, bootrep) %>%
  arrange(antigenf,bootrep, arm) %>%
  # calculate the difference, azithro - placebo
  # in each boostrap replicate
  # note: the factor order is placebo, azithro
  mutate(diff_seroprev = seroprev - lag(seroprev),
         diff_seroprev_minus20 = seroprev_minus20 - lag(seroprev_minus20),
         diff_seroprev_plus20 = seroprev_plus20 - lag(seroprev_plus20)) %>%
  ungroup() %>%
  group_by(antigenf) %>%
  filter(!is.na(diff_seroprev)) %>%
  mutate(diff_seroprev_lb = quantile(diff_seroprev, probs = 0.025),
         diff_seroprev_ub = quantile(diff_seroprev, probs = 0.975),
         diff_seroprev_minus20_lb = quantile(diff_seroprev_minus20, probs = 0.025),
         diff_seroprev_minus20_ub = quantile(diff_seroprev_minus20, probs = 0.975),
         diff_seroprev_plus20_lb = quantile(diff_seroprev_plus20, probs = 0.025),
         diff_seroprev_plus20_ub = quantile(diff_seroprev_plus20, probs = 0.975)
         ) %>%
  select(antigenf,starts_with("diff_")) %>%
  slice(1)
  
mu_serop2 <- mu_serop %>%
  left_join(mu_serop_ci, by = "antigenf")
  
```


```{r cutoff sensitivity seroprev diff figure}
#------------------------------
# plot the difference between
# arms (companion to the means)
#------------------------------

# format the output for plotting
# reshape long
mu_serop_plot_primary <- mu_serop2 %>%
  ungroup() %>%
  group_by(antigenf) %>%
  select(antigenf,diff,diff_lb=diff_seroprev_lb,diff_ub=diff_seroprev_ub) %>%
  mutate(analysis = "Primary analysis")

mu_serop_plot_minus20 <- mu_serop2 %>%
  ungroup() %>%
  group_by(antigenf) %>%
  select(antigenf,diff=diff_minus20,diff_lb=diff_seroprev_minus20_lb,diff_ub=diff_seroprev_minus20_ub) %>%
  mutate(analysis = "MFI cutoff -20%")

mu_serop_plot_plus20 <- mu_serop2 %>%
  ungroup() %>%
  group_by(antigenf) %>%
  select(antigenf,diff=diff_plus20,diff_lb=diff_seroprev_plus20_lb,diff_ub=diff_seroprev_plus20_ub)  %>%
  mutate(analysis = "MFI cutoff +20%")

mu_serop_plot <- mu_serop_plot_primary %>%
  bind_rows(mu_serop_plot_minus20,mu_serop_plot_plus20) %>%
  mutate(analysis = factor(analysis, levels = c("Primary analysis","MFI cutoff -20%","MFI cutoff +20%"))) %>%
  arrange(antigenf,analysis)
mu_serop_plot$pathogen_group <- NA
mu_serop_plot$pathogen_group[grep("P. ",mu_serop_plot$antigenf)] <- "malaria"
mu_serop_plot$pathogen_group[is.na(mu_serop_plot$pathogen_group)] <- "bacteria & protozoa"

# make the figure
pcols <- c("black",cbpal[c(2,7)])
plot_serop_sens <- ggplot(data = mu_serop_plot, aes(x = antigenf, y = diff, color = analysis)) +
  facet_grid(.~pathogen_group, scales = "free_x", space = "free_x") +
  geom_hline(yintercept = 0, color = "gray60") +
  geom_errorbar(aes(ymin = diff_lb, ymax = diff_ub),width = 0.15,alpha=1, position = position_nudge(x = rep(c(-0.2,0,0.2),15))) +
  geom_point(alpha=1, position = position_nudge(x = rep(c(-0.2,0,0.2),15))) +
  scale_color_manual(values = pcols) +
  scale_y_continuous(breaks = seq(-0.2,0.1,by=0.05), labels=sprintf("%1.0f",seq(-0.2,0.1,by=0.05)*100)) +
  labs(x = NULL, y = "difference in seroprevalence (%)\nazithromycin - placebo",tag="a") +
  coord_cartesian(ylim = c(-0.22,0.11)) +
  # theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle=90,hjust=1),
    plot.tag = element_text(face = "bold")
  )

plot_serop_sens

ggsave(here("figures","mordor-ab-diff-seroprev-cutoff-sens.png"),plot_serop_sens,device = "png",width = 7, height = 6)
```


# Force of infection by arm

Estimate serological force of infection through the seroconverion rate. Estimate the seroconversion rate from age-structured seroprevalence as in the primary anlaysis.

```{r FOI  estimates}
#---------------------------------
# for enterics, drop children >24 mo
# (excluding Salmonella)
#---------------------------------
d_foi <- d_serop %>%
  mutate(dropobs = ifelse(antigenf %in% c("Campylobacter p18 or p39","ETEC LTB","Cryptosporidium Cp17 or Cp23","Giardia VSP-3 or VSP-5") & (agem > 24),1,0)) %>%
  filter(dropobs==0) %>%
  select(-dropobs) %>%
  mutate(antigenf = droplevels(antigenf),
         clusteridf = factor(clusterid),
         dummy = 1)

#---------------------------------
# estimate SCRs and HRs for each
# pathogen
#
# repeat for the primary analysis,
# and the two sensitivity analyses
#---------------------------------
set.seed(2342)
foi_ests <- foreach(ab=levels(d_foi$antigenf),.combine=rbind) %do% {
  
    pd <- filter(d_foi, antigenf==ab)
    gfit <- gam(seropos ~ s(agem, bs="cr") + arm +
                s(clusteridf, bs = "re",by=dummy), 
                data=pd,
                family=binomial(link="cloglog"))
    gsum <- summary(gfit)
    # estimate average seroconversion rate by arm
    # use a parametric bootstrap of the posterior VCOV matrix to get inference
    newd0 <- data.frame(agem=c(min(pd$agem),max(pd$agem)), arm = "placebo", dummy=0, clusteridf = "community-1")
    newd1 <- data.frame(agem=c(min(pd$agem),max(pd$agem)), arm = "azithro", dummy=0, clusteridf = "community-1")
    lambda0 <- avgFOI(gfit, newdata=newd0, a1=min(pd$agem),a2=max(pd$agem),nreps=10000)
    lambda1 <- avgFOI(gfit, newdata=newd1, a1=min(pd$agem),a2=max(pd$agem),nreps=10000)
    # estimate HR and 95% CI from the model from the model 
    hr <- as.numeric(exp(gfit$coefficients[2]))
    loghr_se <-as.numeric(gsum$se[2])
    hr_p <- as.numeric(gsum$p.pv[2])
    # return the results
    res <- data.frame(antigenf = ab,
                      lambda0 = lambda0$mufoi,
                      lambda0_se = lambda0$mufoi_se,
                      lambda0_min95 = lambda0$mufoi_lb,
                      lambda0_max95 = lambda0$mufoi_ub,
                      lambda1 = lambda1$mufoi,
                      lambda1_se = lambda1$mufoi_se,
                      lambda1_min95 = lambda1$mufoi_lb,
                      lambda1_max95 = lambda1$mufoi_ub,
                      hr = hr, 
                      loghr_se = loghr_se,
                      hr_min95 = exp(log(hr) - 1.96*loghr_se),
                      hr_max95 = exp(log(hr) + 1.96*loghr_se),
                      hr_p = hr_p
                      )
}

foi_ests_minus20 <- foreach(ab=levels(d_foi$antigenf),.combine=rbind) %do% {
  
    pd <- filter(d_foi, antigenf==ab)
    gfit <- gam(seropos_minus20 ~ s(agem, bs="cr") + arm +
                s(clusteridf, bs = "re",by=dummy), 
                data=pd,
                family=binomial(link="cloglog"))
    gsum <- summary(gfit)
    # estimate average seroconversion rate by arm
    # use a parametric bootstrap of the posterior VCOV matrix to get inference
    newd0 <- data.frame(agem=c(min(pd$agem),max(pd$agem)), arm = "placebo", dummy=0, clusteridf = "community-1")
    newd1 <- data.frame(agem=c(min(pd$agem),max(pd$agem)), arm = "azithro", dummy=0, clusteridf = "community-1")
    lambda0 <- avgFOI(gfit, newdata=newd0, a1=min(pd$agem),a2=max(pd$agem),nreps=10000)
    lambda1 <- avgFOI(gfit, newdata=newd1, a1=min(pd$agem),a2=max(pd$agem),nreps=10000)
    # estimate HR and 95% CI from the model from the model 
    hr <- as.numeric(exp(gfit$coefficients[2]))
    loghr_se <-as.numeric(gsum$se[2])
    hr_p <- as.numeric(gsum$p.pv[2])
    # return the results
    res <- data.frame(antigenf = ab,
                      lambda0 = lambda0$mufoi,
                      lambda0_se = lambda0$mufoi_se,
                      lambda0_min95 = lambda0$mufoi_lb,
                      lambda0_max95 = lambda0$mufoi_ub,
                      lambda1 = lambda1$mufoi,
                      lambda1_se = lambda1$mufoi_se,
                      lambda1_min95 = lambda1$mufoi_lb,
                      lambda1_max95 = lambda1$mufoi_ub,
                      hr = hr, 
                      loghr_se = loghr_se,
                      hr_min95 = exp(log(hr) - 1.96*loghr_se),
                      hr_max95 = exp(log(hr) + 1.96*loghr_se),
                      hr_p = hr_p
                      )
}

foi_ests_plus20 <- foreach(ab=levels(d_foi$antigenf),.combine=rbind) %do% {
  
    pd <- filter(d_foi, antigenf==ab)
    gfit <- gam(seropos_plus20 ~ s(agem, bs="cr") + arm +
                s(clusteridf, bs = "re",by=dummy), 
                data=pd,
                family=binomial(link="cloglog"))
    gsum <- summary(gfit)
    # estimate average seroconversion rate by arm
    # use a parametric bootstrap of the posterior VCOV matrix to get inference
    newd0 <- data.frame(agem=c(min(pd$agem),max(pd$agem)), arm = "placebo", dummy=0, clusteridf = "community-1")
    newd1 <- data.frame(agem=c(min(pd$agem),max(pd$agem)), arm = "azithro", dummy=0, clusteridf = "community-1")
    lambda0 <- avgFOI(gfit, newdata=newd0, a1=min(pd$agem),a2=max(pd$agem),nreps=10000)
    lambda1 <- avgFOI(gfit, newdata=newd1, a1=min(pd$agem),a2=max(pd$agem),nreps=10000)
    # estimate HR and 95% CI from the model from the model 
    hr <- as.numeric(exp(gfit$coefficients[2]))
    loghr_se <-as.numeric(gsum$se[2])
    hr_p <- as.numeric(gsum$p.pv[2])
    # return the results
    res <- data.frame(antigenf = ab,
                      lambda0 = lambda0$mufoi,
                      lambda0_se = lambda0$mufoi_se,
                      lambda0_min95 = lambda0$mufoi_lb,
                      lambda0_max95 = lambda0$mufoi_ub,
                      lambda1 = lambda1$mufoi,
                      lambda1_se = lambda1$mufoi_se,
                      lambda1_min95 = lambda1$mufoi_lb,
                      lambda1_max95 = lambda1$mufoi_ub,
                      hr = hr, 
                      loghr_se = loghr_se,
                      hr_min95 = exp(log(hr) - 1.96*loghr_se),
                      hr_max95 = exp(log(hr) + 1.96*loghr_se),
                      hr_p = hr_p
                      )
}

#---------------------------------
# combine the three sets of estimates
# into a single results data frame
#---------------------------------
foi_ests_all <- foi_ests %>%
  mutate(analysis = "Primary analysis") %>%
  bind_rows(foi_ests_minus20) %>%
  mutate(analysis = ifelse(is.na(analysis),"MFI cutoff -20%",analysis)) %>%
  bind_rows(foi_ests_plus20) %>%
  mutate(analysis = ifelse(is.na(analysis),"MFI cutoff +20%",analysis)) %>%
  mutate(analysis = factor(analysis, levels=c("Primary analysis","MFI cutoff -20%","MFI cutoff +20%")))
```


## Sensitivity analysis HR figure


```{r sensitivity scr hazard ratio figure}
#------------------------------
# plot the hazard ratio between
# arms (companion to the means)
#------------------------------

# format the output for plotting
mu_scr_plot <- foi_ests_all %>%
  ungroup() %>%
  arrange(antigenf,analysis)
mu_scr_plot$pathogen_group = "bacteria & protozoa"
mu_scr_plot$pathogen_group[substr(as.character(mu_scr_plot$antigenf),1,3)=="P. "] <- "malaria"

# make the figure
pcols <- c("black",cbpal[c(2,7)])
plot_hr_sens <- ggplot(data = mu_scr_plot, aes(x = antigenf, y = hr, color = analysis)) +
  facet_grid(.~pathogen_group, scales = "free_x", space = "free_x") +
  geom_hline(yintercept = 1, color = "gray60") +
  geom_errorbar(aes(ymin = hr_min95, ymax = hr_max95),width = 0.15,alpha=1, position = position_nudge(x = rep(c(-0.2,0,0.2),15))) +
  geom_point(alpha=1, position = position_nudge(x = rep(c(-0.2,0,0.2),15))) +
  scale_color_manual(values = pcols) +
  scale_y_continuous(breaks = c(0.5,0.8,1,1.2,1.5,2),trans = "log") +
  labs(x = NULL, y = "hazard ratio \nazithromycin / placebo",tag="b") +
  coord_cartesian(ylim = c(0.4,2)) +
  # theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle=90,hjust=1),
    plot.tag = element_text(face = "bold")
  )

plot_hr_sens

ggsave(here("figures","mordor-ab-hr-cutoff-sens.png"),plot_hr_sens,device = "png",width = 7, height = 6)
```


# Save estimates

```{r save hr estimates}
write_csv(mu_serop_plot,here("data","mordor-ab-cross-sectional-serop-ests-cutoff-sensitivity-analysis.csv"))
write_csv(mu_scr_plot,here("data","mordor-ab-cross-sectional-hr-ests-cutoff-sensitivity-analysis.csv"))
```

# Session Info
```{r session info}
sessionInfo()
```


