---
title: "Effect of biannual azithromycin distribution on antibody responses to malaria, bacterial, and protozoan pathogens in Niger"
subtitle: "Leave-one-out jackknife sensitivity analysis"
author: "Ben Arnold ben.arnold@ucsf.edu"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: haddock
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Summary

Repeat the primary analysis that compares azithromycin and placebo groups by seropositivity and seroconversion rates, systematically excluding one community in turn and re-estimating the effect. Use the mean over replicates for a jackknife estimate of the bias.

# Preamble

```{r preamble, message=FALSE}
library(here)
here()
#----------------------------
# source the config file
#----------------------------
source(here("R","mordor-ab-Config.R"))

#----------------------------
# source the functions file
# includes reused functions
#----------------------------
source(here("R","mordor-ab-Functions.R"))
```




# Load the antibody data

```{r load antibody data}
#-------------------------------
# load the formatted analysis
# data created by
# mordor-ab-format-data.Rmd
#-------------------------------
d <- read_rds(here("data","mordor-ab-analysis-public.rds"))

#-------------------------------
# drop "_public" suffix from IDs
#-------------------------------
d <- d %>%
  rename(clusterid = clusterid_public, childid = childid_public)

#-------------------------------
# drop GST negative control values
# drop trachoma and strongy because
# no transmission in this population
#-------------------------------
d1 <- d %>%
  filter(!antigen %in% c("gst","pgp3","ct694","nie")) %>%
  mutate(antigenf = factor(antigenf))

#-------------------------------
# drop cholera CTB because so 
# highly cross-reactive with ETEC LTB
#-------------------------------
d1 <- d1 %>%
  filter( !antigen %in% c("ctb") ) %>%
  mutate(antigenf = factor(antigenf))

#-------------------------------
# drop baseline measurements
#-------------------------------
d2 <- d1 %>%
  filter(phase > 0)

#-------------------------------
# age restrictions based on 
# age-antibody curves
# to exclude ages with evidence
# for maternal IgG contributions
#
# Bacteria and protozoa: 6 mo - 59 mo
#
# ETEC LTB: 6 mo - 24 mo (no heterogeneity at older ages)
#
# Malaria antigens: 12 - 59 mo
#-------------------------------
d2 <- d2 %>%
  filter(agem >= 6) %>%
  mutate(agedrop = case_when(
    antigen %in% c("etec") & (agem > 24) ~ 1,
    antigen %in% c("hrp2","csp","glurp","lsa1","ama","pfmsp1","pvmsp1","pomsp1","pmmsp1") & (agem < 12) ~ 1,
    TRUE ~ 0
  )
         ) %>%
  filter(agedrop == 0)

```

# Create composite seropositivity indicators

For enteric pathogens with multiple antigens (Cryptosporidium, Giardia, Salmonella, Campylobacter), create a series of seropositivity indicators that indicate if a child was positive to either antigen.

Then bind those to the original data frame to create a seropositivity dataset for analyses.

```{r composite seropositivity indicators}
d_enteric_comps <- d2 %>%
  filter(pathogen %in% c("Cryptosporidium","Giardia","Salmonella","Campylobacter")) %>%
  group_by(pathogen,childid,phase) %>%
  mutate(seropos = max(seropos)) %>%
  slice(1) %>%
  mutate(antigenf = case_when(
    antigen %in% c("cp17","cp23") ~ "Cryptosporidium Cp17 or Cp23",
    antigen %in% c("vsp3","vsp5") ~ "Giardia VSP-3 or VSP-5",
    antigen %in% c("salb","sald") ~ "Salmonella groups B or D LPS",
    antigen %in% c("p18","p39") ~ "Campylobacter p18 or p39"
  )) %>%
  select(clusterid,phase,arm,childid,agem,sex,dbsid,pathogen,antigenf,seropos)


```

```{r create seropositivity dataset, warning=FALSE}
d_serop <- d2 %>%
  # drop pathogens w/ composite indicators
  filter(!pathogen %in% c("Cryptosporidium","Giardia","Salmonella","Campylobacter")) %>%
  # add composite indicators
  bind_rows(d_enteric_comps) %>%
  # create new factor levels that includes composite indicators
  mutate(antigenf = factor(antigenf,levels = c(
    "Campylobacter p18 or p39",
    "ETEC LTB",
    "Salmonella groups B or D LPS",
    "Cryptosporidium Cp17 or Cp23",
    "Giardia VSP-3 or VSP-5",
   levels(d2$antigenf)[10:19]) )
         ) %>%
  # drop extra variables (MFI doesn't correspond to composite seropos so drop)
  dplyr::select(-mfi,-logmfi,-agedrop)

```




# Seroprevalence

Estimate means by arm in the full sample, then re-estimate means with leave-one-out at the community level (30 replicates)

```{r seroprev by arm}

#-----------------------------
# tally number of observations
# by antigen (including composites)
#-----------------------------
n_serop <- d_serop %>%
  group_by(pathogen,antigenf,arm) %>%
  mutate(nomiss = ifelse(!is.na(seropos),1,0)) %>%
  summarize(nobs = sum(nomiss), .groups = "keep") %>%
  pivot_wider(id_cols = c(pathogen, antigenf), names_from = "arm", values_from = nobs) %>%
  rename(placebo_n = placebo,
         azithro_n = azithro)

# print Ns for reporting in figure legends
(n_serop)

#-----------------------------
# estimate seroprevalence by arm
# and difference in seroprevalence
#-----------------------------
mu_serop <- d_serop %>%
  group_by(pathogen,antigenf,arm) %>%
  summarize(mean_serop = mean(seropos), .groups = "keep") %>%
  pivot_wider(id_cols = c(pathogen, antigenf), names_from = "arm", values_from = mean_serop) %>%
  mutate(diff = azithro - placebo)

#-----------------------------
# merge means onto the Ns
#-----------------------------
mu_serop <- left_join(n_serop, mu_serop, by=c("pathogen","antigenf"))
```

```{r bootstrap mean seropos diff}
#-----------------------------
# bootstrap resample communities
# with replacement to estimate
# 95% CIs for difference in seroprev
#-----------------------------
nreps <- 1000
ncl <- length(unique(d_serop$clusterid))
bsamp <- matrix(sample(unique(d_serop$clusterid),size = ncl*nreps, replace = TRUE),nrow = ncl,ncol = nreps)
d_serop_boot <- d_serop %>% ungroup %>% select(clusterid,antigenf,arm,seropos)
mu_serop_boot <- foreach(bi = 1:nreps, .combine = rbind) %dopar% {
  set.seed(bi)
  di <- left_join(data.frame(clusterid = bsamp[,bi], stringsAsFactors = FALSE),
                  d_serop_boot, by = "clusterid")
  di %>%
  group_by(antigenf,arm) %>%
  summarize(mean_serop = mean(seropos), .groups = "keep") %>%
  pivot_wider(id_cols = "antigenf", names_from = "arm", values_from = mean_serop) %>%
  mutate(diff = azithro - placebo, 
         bootrep = bi)

}
```

```{r mean seroprev bootstrap cis}
#-----------------------------
# calculate percentile 95%
# CIs from the bootstrap
# distribution merge to estimates
#-----------------------------
mu_serop_ci <- mu_serop_boot %>%
  group_by(antigenf) %>%
  mutate(mean_p_lb = quantile(placebo, probs = 0.025),
         mean_p_ub = quantile(placebo, probs = 0.975),
         mean_a_lb = quantile(azithro, probs = 0.025),
         mean_a_ub = quantile(azithro, probs = 0.975),
         diff_bmean = mean(diff),
         diff_bse = sd(diff),
         diff_lb = quantile(diff, probs = 0.025),
         diff_ub = quantile(diff, probs = 0.975)) %>%
  select(antigenf,starts_with("mean_"), starts_with("diff_")) %>%
  slice(1)
  
mu_serop2 <- mu_serop %>%
  left_join(mu_serop_ci, by = "antigenf")
  
```


```{r loo seroprevalence}
#-----------------------------
# loop over the 30 communities
# and re-estimate difference
# in seroprevalence
#-----------------------------
loo_serop <- foreach(comi = unique(d_serop$clusterid), .combine = rbind) %do% {
  
  # estimate seroprevalence by arm
  # and difference in seroprevalence
  mui <- d_serop %>% 
    filter(clusterid != comi) %>%
    group_by(pathogen,antigenf,arm) %>%
    summarize(mean_serop = mean(seropos), .groups = "keep") %>%
    pivot_wider(id_cols = c(pathogen, antigenf), names_from = "arm", values_from = mean_serop) %>%
    mutate(diff = azithro - placebo,
           loo = comi)
}

#-----------------------------
# calculate the mean over
# replicates
#-----------------------------
loo_mu <- loo_serop %>%
  group_by(pathogen, antigenf) %>%
  summarize(azithro = mean(azithro),
            placebo = mean(placebo),
            diff = mean(diff),
            .groups = "keep")
loo_mu$group = "bacteria & protozoa"
loo_mu$group[grep("P. ",loo_mu$pathogen)] <- "malaria"

#-----------------------------
# jackknife estimation of the
# bias: (n-1)(theta_loo - theta)
# https://en.wikipedia.org/wiki/Jackknife_resampling
#-----------------------------
loo_mu2 <- loo_mu %>%
  rename(diff_loo = diff ) %>%
  left_join(mu_serop2 %>% select(pathogen, antigenf, diff, diff_lb, diff_ub), by = c("pathogen","antigenf")) %>%
  mutate(jkn_bias = 29*(diff_loo - diff)*100,
         jkn_bias_lab = sprintf("%1.2f",jkn_bias))


```

```{r seroprev loo summary, warning=FALSE}
set.seed(1347981)
loo_serop_sum <- loo_serop
loo_serop_sum$group = "bacteria & protozoa"
loo_serop_sum$group[grep("P. ",loo_serop_sum$pathogen)] <- "malaria"

loo_serop_plot <- ggplot(data = loo_serop_sum, aes(x = antigenf, y = diff)) +
  facet_grid(.~group, scales = "free_x", space = "free_x") +
  geom_hline(yintercept = 0, color = "gray60") +
  geom_jitter(width = 0.25, alpha = 0.4, color = cbpal[7]) +
  geom_point(data=loo_mu2, aes(x = antigenf, y = diff_loo), pch = 1, size = 3,color = "black") +
  geom_point(data=loo_mu2, aes(x = antigenf, y = diff),pch=3, size=5) +
  geom_errorbar(data=loo_mu2, aes(x = antigenf, ymin = diff_lb, ymax = diff_ub),width = 0) +
  geom_text(data=loo_mu2, aes(x = antigenf, y = -0.24, label = jkn_bias_lab),pch=1, size=3) +
  scale_y_continuous(breaks = seq(-0.25,0.1,by=0.05), labels=sprintf("%1.0f",seq(-0.25,0.1,by=0.05)*100)) +
  coord_cartesian(ylim=c(-0.25,0.11)) +
  labs(x = "",y = "Difference in seroprevalence (%)\nazithromycin - placebo", tag = "a") +
  theme(
    axis.text.x = element_text(angle=90,hjust=1),
    plot.tag = element_text(face = "bold")
  )

loo_serop_plot

ggsave(here("figures","mordor-ab-diff-seroprev-loo-sens.pdf"),loo_serop_plot,device = "pdf",width = 7, height = 6)

```
Leave-one-out sensitivity analyses that excluded each of the 30 study communities and re-estimated the difference in seroprevalence between groups and the relative hazard of seroconversion. Light points mark the 30 leave-one-out replicates, black circles mark the mean over leave-one-out replicates. Horizontal lines mark the mean in the full sample and vertical lines the 95% confidence interval. Jackknife estimates of the bias are printed at the bottom of each panel, above the pathogen/antigen labels. (**a**) Difference in seroprevalence between groups. Jackknife estimates of the bias are on the percentage point scale, all <0.03%. 

# Force of infection by arm

Estimate serological force of infection through the seroconverion rate. Estimate the seroconversion rate from age-structured seroprevalence as in the primary anlaysis.


```{r full sample HR estimates}
#---------------------------------
# for enterics, drop children >24 mo
# (excluding Salmonella)
#---------------------------------
d_foi <- d_serop %>%
  mutate(dropobs = ifelse(antigenf %in% c("Campylobacter p18 or p39","ETEC LTB","Cryptosporidium Cp17 or Cp23","Giardia VSP-3 or VSP-5") & (agem > 24),1,0)) %>%
  filter(dropobs==0) %>%
  select(-dropobs) %>%
  mutate(antigenf = droplevels(antigenf),
         clusteridf = factor(clusterid),
         dummy = 1)

foi_ests <- foreach(ab=levels(d_foi$antigenf),.combine=rbind) %do% {
  
    pd <- filter(d_foi, antigenf==ab)
    gfit <- gam(seropos ~ s(agem, bs="cr") + arm +
                s(clusteridf, bs = "re",by=dummy), 
                data=pd,
                family=binomial(link="cloglog"))
    gsum <- summary(gfit)
    # estimate HR and 95% CI from the model from the model 
    hr <- as.numeric(exp(gfit$coefficients[2]))
    loghr_se <-as.numeric(gsum$se[2])
    hr_p <- as.numeric(gsum$p.pv[2])
    # return the results
    res <- data.frame(antigenf = ab,
                      hr = hr, 
                      loghr_se = loghr_se,
                      hr_min95 = exp(log(hr) - 1.96*loghr_se),
                      hr_max95 = exp(log(hr) + 1.96*loghr_se),
                      hr_p = hr_p
                      )
}
```

```{r loo HR estimates}
#-----------------------------
# for each antigen:
# loop over the 30 communities
# and re-estimate HR
# from age-structured seroprev
#-----------------------------
loo_loghr <- foreach(ab=levels(d_foi$antigenf),.combine=rbind) %:%
  foreach(comi = unique(d_serop$clusterid), .combine = rbind) %do% {
    
    # exclude the loo cluster
    di <- d_foi %>% filter(clusterid != comi & antigenf == ab)
    
    # estimate log HR from the model 
    gfit <- gam(seropos ~ s(agem, bs="cr") + arm +
                s(clusteridf, bs = "re",by=dummy), 
                data=di,
                family=binomial(link="cloglog"))

    # return the results
    res <- data.frame(loo = comi,
                      antigenf = ab,
                      loghr = as.numeric(gfit$coefficients[2])
                      )
   res
}

#-----------------------------
# calculate the mean log(HR)
# over replicates
#-----------------------------
loo_loghr_mu <- loo_loghr %>%
  group_by(antigenf) %>%
  summarize(loghr = mean(loghr),.groups = "keep") %>%
  mutate(antigenf = factor(antigenf, levels = levels(d_foi$antigenf))) %>%
  arrange(antigenf)
loo_loghr_mu$group = "bacteria & protozoa"
loo_loghr_mu$group[grep("P. ",loo_loghr_mu$antigenf)] <- "malaria"

#-----------------------------
# jackknife estimation of the
# bias: (n-1)(theta_loo - theta)
# https://en.wikipedia.org/wiki/Jackknife_resampling
#-----------------------------
loo_loghr_mu2 <- loo_loghr_mu %>%
  rename(loghr_loo = loghr ) %>%
  left_join(foi_ests %>% select(antigenf, hr, hr_min95, hr_max95), by = c("antigenf")) %>%
  mutate(hr_loo = exp(loghr_loo),
         jkn_bias = 29*(loghr_loo - log(hr)),
         jkn_bias_lab = sprintf("%1.2f",jkn_bias))


```


```{r HR loo summary, warning = FALSE}
set.seed(3274)
loo_loghr_sum <- loo_loghr %>%
  mutate(hr = exp(loghr)) %>%
  mutate(antigenf = factor(antigenf, levels = levels(d_foi$antigenf)))
loo_loghr_sum$group = "bacteria & protozoa"
loo_loghr_sum$group[grep("P. ",loo_loghr_sum$antigenf)] <- "malaria"
  

loo_hr_plot <- ggplot(data = loo_loghr_sum , aes(x = antigenf, y = hr)) +
  facet_grid(.~group, scales = "free_x", space = "free_x") +
  geom_hline(yintercept = 1, color = "gray60") +
  geom_jitter(width = 0.25, alpha = 0.4, color = cbpal[7]) +
  geom_point(data=loo_loghr_mu2, pch = 3, size = 5,color = "black") +
  geom_point(data=loo_loghr_mu2, aes(x = antigenf, y = hr_loo),pch=1, size=3) +
  geom_errorbar(data=loo_loghr_mu2, aes(x = antigenf, ymin = hr_min95, ymax = hr_max95),width = 0) +
  geom_text(data=loo_loghr_mu2, aes(x = antigenf, y = 0.35, label = jkn_bias_lab),pch=1, size=3) +
  scale_y_continuous(breaks = c(0.5,0.8,1,1.2,1.5,2),trans = "log") +
  coord_cartesian(ylim=c(0.3,2)) +
  labs(x = "",y = "Hazard ratio of seroconversion\nazithromycin / placebo", tag = "b") +
  theme(
    axis.text.x = element_text(angle=90,hjust=1),
    plot.tag = element_text(face = "bold")
  )

loo_hr_plot

ggsave(here("figures","mordor-ab-hr-loo-sens.pdf"),loo_hr_plot,device = "pdf",width = 7, height = 6)

```

Leave-one-out sensitivity analyses that excluded each of the 30 study communities and re-estimated the difference in seroprevalence between groups and the relative hazard of seroconversion. Light points mark the 30 leave-one-out replicates, black circles mark the mean over leave-one-out replicates. Horizontal lines mark the mean in the full sample and vertical lines the 95% confidence interval. Jackknife estimates of the bias are printed at the bottom of each panel, above the pathogen/antigen labels.(**b**) Hazard ratio of seroconversion estimated from age-structured seroprevalence. Jackknife estimates of the bias are on the log(hazard ratio) scale. 


# Session Info
```{r session info}
sessionInfo()
```


