---
title: "Analysis of multiplex antibody endpoints in the MORDOR Niger trial"
subtitle: "Assess correlation between ETEC LTB and cholera CTB"
author: "Ben Arnold ben.arnold@ucsf.edu"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: haddock
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Summary

Antibody responses to ETEC labile toxin B subnit (LTB) and cholera toxin B subunit are known to be cross-reactive. Confirm this is the case in these measurements. Based on the results, exclude cholera CTB from the main results because it's difficult to interpret without additional antibody markers.

Arnold BF, et al. Enteropathogen antibody dynamics and force of infection among children in low-resource settings. _eLife_ 8, (2019). https://pubmed.ncbi.nlm.nih.gov/31424386/

Smith NW, Sack RB. 1973. Immunologic cross-reactions of enterotoxins from Escherichia coli and Vibrio cholerae. _Journal of Infectious Diseases_ 127:164–170. DOI: https://doi.org/10.1093/infdis/127.2.164, PMID: 4570128

# Preamble

```{r preamble}
library(here)
here()
#----------------------------
# source the config file
#----------------------------
source(here("R","mordor-ab-Config.R"))

#----------------------------
# source the functions file
# includes reused functions
#----------------------------
source(here("R","mordor-ab-Functions.R"))
```



# Load the antibody data

```{r load antibody data}
#-------------------------------
# load the formatted analysis
# data created by
# mordor-ab-format-data.Rmd
#-------------------------------
d <- read_rds(here("data","mordor-ab-analysis.rds"))

#-------------------------------
# limit to ETEC and cholera
#-------------------------------
d1 <- d %>%
  filter(antigen %in% c("etec","ctb")) %>%
  mutate(antigenf = factor(antigenf))



#-------------------------------
# reshape the data wide
# to make it easier to plot
#-------------------------------
dmfi <- d1 %>%
  ungroup() %>%
  select(childid,phase,agem,antigen,logmfi) %>%
  pivot_wider(id_cols = c("childid","phase","agem"), names_from = antigen, values_from = logmfi)

```


# Correlation between ETEC LTB and cholera CTB

```{r ETEC v CTB figure}

# estimate Spearman rank correlation for printing
sp_corr <- cor(dmfi$etec,dmfi$ctb, method = "spearman")
cor_lab <- paste(sprintf("%1.2f",sp_corr))

# custom log labels
log10labs <- c( 
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

plot_etec_ctb <- ggplot(data=dmfi, aes(x = ctb, y = etec)) +
  geom_point(alpha = 0.2, size = 0.3, color = vircols[5]) +
  geom_smooth(color = "black",lwd=0.75,se=FALSE, method = "loess") +
  annotate("text",x=1.1,y=4.25,label = paste("rho ==",deparse(cor_lab)),parse=TRUE, hjust =0) +
  scale_x_continuous(breaks=1:4,labels = log10labs) +
  scale_y_continuous(breaks=1:4,labels = log10labs) +
  labs(x = "Cholera CTB  IgG Luminex Response (MFI-bg)", y = "ETEC LTB IgG Luminex Response (MFI-bg)") +
  theme_minimal()

plot_etec_ctb

ggsave(filename = here("figures","mordor-ab-etec-v-ctb.png"),plot_etec_ctb, device = "png",width=4, height = 4)

```


# Session Info
```{r session info}
sessionInfo()
```


