---
title: "Analysis of multiplex antibody endpoints in the MORDOR Niger trial"
subtitle: "CONSORT participant flow"
author: "Ben Arnold ben.arnold@ucsf.edu"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: haddock
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Summary

Estimate the number of children measured in each round to summarize in a CONSORT flow diagram.

# Preamble

```{r preamble}
library(here)
here()
#----------------------------
# source the config file
#----------------------------
source(here("R","mordor-ab-Config.R"))

#----------------------------
# source the functions file
# includes reused functions
#----------------------------
source(here("R","mordor-ab-Functions.R"))
```



# Load the antibody data

```{r load antibody data}
#-------------------------------
# load the formatted analysis
# data created by
# mordor-ab-format-data.Rmd
#-------------------------------
d <- read_rds(here("data","mordor-ab-analysis.rds"))

#-------------------------------
# drop the trachoma, strongyloides, 
# and GST negative control values
#-------------------------------
d1 <- d %>%
  filter(!antigen %in% c("pgp3","ct694","nie","gst")) %>%
  mutate(antigenf = factor(antigenf))


#-------------------------------
# age restrictions based on 
# age-antibody curves
# to exclude ages without any
# variation in antibody levels
# (esp. enterics, which become saturated)
#
# Enterics: 6 mo - 24
# (except for salmonella, strep A)
#
# Malaria antigens: 12 - 59 mo
#-------------------------------
d2 <- d1 %>%
  filter(agem >= 6) %>%
  mutate(agedrop = case_when(
    antigen %in% c("cp17","cp23","vsp3","vsp5","p18","p39","etec","ctb") & (agem > 24) ~ 1,
    antigen %in% c("hrp2","csp","glurp","lsa1","ama","pfmsp1","pvmsp1","pomsp1","pmmsp1") & (agem < 12) ~ 1,
    TRUE ~ 0
  )
         )

```

# Summarize the number of measurements by phase and arm

## Total measurements by phase and arm (and overall)

Total samples by phase and arm

```{r cross tab of n by arm}
#-----------------------------
# summarize total samples
# analyzed for IgG responses
# by phase and arm
#-----------------------------
d_n <- d1 %>%
  filter(antigen=="vsp3" & !is.na(logmfi))

dim(d_n)
table(d_n$phase)
table(d_n$arm)
table(d_n$phase,d_n$arm)
```

## Ages 6-24 months (enterics)

Number of samples by phase and arm

```{r cross tab of n by arm 6 to 24 months}
#-----------------------------
# summarize samples among 
# children ages 6-24 months
#-----------------------------
d_n2 <- d2 %>%
  filter(antigen == "vsp3" & !is.na(logmfi) & agedrop == 0) 

dim(d_n2)
table(d_n2$phase)
table(d_n2$arm)
table(d_n2$phase,d_n2$arm)
```

Total samples by arm, post-baseline

```{r total samples 6 to 24 months post-treatment}
#-----------------------------
# summarize totals post-treatment
#-----------------------------
dim(d_n2 %>% filter(phase>0))
table(d_n2$arm[d_n2$phase>0])

```

## Ages 12-59 months (malaria)

Number of samples by phase and arm
```{r cross tab of n by arm 12 to 59 months}
#-----------------------------
# summarize samples among 
# children ages 6-24 months
#-----------------------------
d_n3 <- d2 %>%
  filter(antigen == "pfmsp1" & !is.na(logmfi) & agedrop == 0) 

dim(d_n3)
table(d_n3$phase)
table(d_n3$arm)
table(d_n3$phase,d_n3$arm)
```

Total samples by arm, post-baseline
```{r total samples 12 to 59 months post-treatment}
#-----------------------------
# summarize totals post-treatment
#-----------------------------
dim(d_n3 %>% filter(phase>0))
table(d_n3$arm[d_n3$phase>0])

```

# Summarize number of children

## Identify total number of children enrolled

Start with total enrolled

```{r summarize total children enrolled}
#------------------------------------
# Identify unique children -- Total
#------------------------------------
dchild <- d1 %>%
  ungroup() %>%
  group_by(childid) %>%
  slice(1)

dim(dchild)
table(dchild$arm)

```

Then identify number excluded because of <6 months (enterics) 

```{r summarize children excluded under 6 months}
#------------------------------------
# identify children excluded < 6 months
#------------------------------------
dchild6 <- d1 %>%
  filter(agem<6 & antigen=="vsp3") %>%
  group_by(childid) %>%
  slice(1)

dim(dchild6)
table(dchild6$arm)

```

Identify number excluded because of <12 months (malaria) 

```{r summarize children excluded under 12 months}
#------------------------------------
# identify children excluded < 12 months
#------------------------------------
dchild12 <- d1 %>%
  filter(agem<12 & antigen=="ama") %>%
  group_by(childid) %>%
  slice(1)

dim(dchild12)
table(dchild12$arm)

```

Identify number excluded because >24 months (enterics) 

```{r summarize children excluded over 24 months}
#------------------------------------
# identify children excluded >24 months
#------------------------------------
dchild24 <- d1 %>%
  filter(agem>24 & antigen=="vsp3") %>%
  group_by(childid) %>%
  slice(1)

dim(dchild24)
table(dchild24$arm)

```

# Summarize treatment coverage by phase and arm

Read in treatement data

```{r read in treatment data}
dtr <- read_rds(here("data","mordor-ab-treatment.rds"))

```

## Summarize children per cluster

For each phase, summarize the mean number of children per cluster and the SD of cluster means

```{r mean and sd children per cluster}
#-----------------------------
# sum the number of children
# by cluster and by phase
#-----------------------------
dnc <- dtr %>%
  group_by(phase,clusterid,arm) %>%
  summarize(nchildren = n(), .groups = "keep") %>%
  slice(1)

dnc_tab <- dnc %>%
  ungroup() %>%
  group_by(arm,phase) %>%
  summarize(mean_children = mean(nchildren),
            sd_children = sd(nchildren),
            .groups = "keep") %>%
  arrange(arm,phase)

knitr::kable(dnc_tab,
             digits = 0) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")

```

## Summarize treatment coverage

For each phase, summarize the average treatment coverage and the SD of cluster averages

```{r mean and sd treatment coverage}
#-----------------------------
# sum the number of eligible
# children and treated children
# per cluster
#-----------------------------
dcov <- dtr %>%
  group_by(phase,clusterid,arm) %>%
  summarize(trcoverage = mean(treated),
            .groups = "keep") %>%
  slice(1) 

# there is no treatment coverage data in community C51 in phase 6
# (consistent w/ the morbidity monitoring data)
# remove that community from the estimates

dcov_tab <- dcov %>%
  ungroup() %>%
  # remove community C51 in phase 6
  filter(!(clusterid=="C51" & phase==6)) %>%
  group_by(arm,phase) %>%
  summarize(mean_coverage = mean(trcoverage)*100,
            sd_coverage = sd(trcoverage)*100,
            .groups = "keep") %>%
  arrange(arm,phase)

knitr::kable(dcov_tab,
             digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = "striped")

```


# Session Info
```{r session info}
sessionInfo()
```


