---
title: "Analysis of multiplex antibody endpoints in the MORDOR Niger trial"
subtitle: "Format GPS and climate data for analysis"
author: "Ben Arnold ben.arnold@ucsf.edu"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: haddock
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Preamble

```{r preamble}
library(here)
here()
#----------------------------
# source the config file
#----------------------------
source(here("R","mordor-ab-Config.R"))


#----------------------------
# spatial packages
# used only for this script
# (it's a pain to load them
# for all scripts b/c raster
# conflicts with dplyr::select)
#----------------------------
library(raster)
library(ncdf4)

```

# Load cluster list from morbidity data

Read in the final morbidity dataset shared by Jermey Keenan in December 2019. 

```{r format morbidity}
#-------------------------------
# read in the full morbidity file
#-------------------------------
md <- read_csv("~/Box sync/mordor/mordor data/data/final-morbidity/MORDOR-Morbidity-merged-20Mar2019.csv",
                      col_types = cols(code_bspot_ = col_character(),
                                       arm = col_character(),
                                       txlabelgwu = col_character(),
                                       txdate = col_character(),
                                       dosex = col_character()))


#-------------------------------
# The code below follows the
# approach used in the main
# data formatting script. But,
# NOTE: this is slightly fragile
# in that it creates a new 
# cluster ID that should be the
# same as the main data processing
# step (confirmed through inspection)
# but is not necessarily as robust
# as retaining the cluster name
# however, we have dropped the
# cluster name from the main 
# data set to protect confidentiality
#-------------------------------

  
#-------------------------------
# filter out observations with
# gwunotes_ not missing (indicates a problem observation)
#
# create a community ID variable
# that isn't based on names
#
# for each community and phase
# create a median treatment
# date
#-------------------------------
md2 <- md %>% 
  dplyr::select(gwu,gwunotes_) %>%
  filter(is.na(gwunotes_)) %>%
  dplyr::select(-gwunotes_) %>%
  group_by(gwu) %>%
  slice(1) %>%
  mutate(cluster_name = str_to_upper(gwu))
  
md2$clusterid <- paste0("C",group_indices(md2,cluster_name)) 

```

# Load MORDOR GPS

Load GPS centroids for MORDOR morbidity clusters

```{r load mordor gps}
#-------------------------------
# Load MORDOR morbidity cluster
# GPS coordinates
#-------------------------------
d_gps <- read_csv(here("data","mordor-morbidity-cluster-gps.csv")) %>%
  mutate(cluster_name = str_to_upper(gwu))

#-------------------------------
# merge the cluster IDs
# onto the GPS coordinates
#-------------------------------
d_gps2 <- d_gps %>%
  left_join(md2, by = "cluster_name") %>%
  dplyr::select(clusterid,cluster_name,longitude,latitude)

```


# Load Climate Data

Monthly precipitation data, global extent

CRU TS4.04: Climatic Research Unit (CRU) Time-Series (TS) version 4.04 of high-resolution gridded data of month-by-month variation in climate (Jan. 1901- Dec. 2019)

Harris, I., Osborn, T. J., Jones, P. & Lister, D. Version 4 of the CRU TS monthly high-resolution gridded multivariate climate dataset. Sci Data 7, 109 (2020).
https://www.nature.com/articles/s41597-020-0453-3
https://www.ncbi.nlm.nih.gov/pubmed/32246091

http://data.ceda.ac.uk/badc/cru/data/cru_ts/cru_ts_4.04/data/pre
https://www.benjaminbell.co.uk/2018/01/getting-climate-data.html

```{r load CRU data}
#-------------------------------
# Load the CRU TS precipitation 
# dataset into R 
#-------------------------------
d_pre <- brick(here("data/cru_ts4_04_2011-2019","cru_ts4.04.2011.2019.pre.dat.nc"), varname="pre")

```

# Spatial merge of CRU to clusters

```{r spatial merge of precip data}
#-------------------------------
# Spatial merge of CRU time series 
# to the 30 MORDOR morbidity clusters
#-------------------------------
mordor_pre <- data.frame(clusterid=d_gps2$clusterid,
                         cluster_name=d_gps2$cluster_name,
                         extract(d_pre, d_gps2[c("longitude","latitude")])
)

#-------------------------------
# reshape the data from wide
# to long format
#-------------------------------
d_pre2 <- mordor_pre %>%
  group_by(clusterid,cluster_name) %>%
  pivot_longer(cols = starts_with("X"),names_to = "rawdate", names_prefix = "X", values_to = "pre")

#-------------------------------
# format the dates
#-------------------------------
d_pre3 <- d_pre2 %>%
  mutate(pdate = as_date(rawdate))

```

# Plot precipitation time series 

Plot the precipitation time series for the 30 communities

All communities share the same overall precipitation patterns.  Therefore, we can be confident in defining wet seasons consistently across communities.

```{r plot precip time series}
prets_plot <- ggplot(data = d_pre3, aes(x = pdate, y = pre, color = clusterid)) +
  geom_line(alpha = 0.3) +
  labs(y = "monthly precipitation (mm)", x = "date")

prets_plot


```

# Identify the wet season

Create indicators of "wet season", defined as >4 mm rainfall in each month. A 4 mm cutoff was chosen based on visual inspection of the classification across years in the 2011 - 2019 time series (below).

```{r code wet season}
d_pre4 <- d_pre3 %>%
  mutate(season = ifelse(pre>4, "wet season","dry season"),
         season = factor(season))


```

Plot of precipitation with seasons indicated

```{r plot precip time series coded as season}
prets_plot2 <- ggplot(data = d_pre4 %>% filter(clusterid=="C7"), aes(x = pdate, y = pre)) +
  geom_polygon(alpha = 0.2) +
  geom_point(aes(color = season), alpha = 0.4) +
  scale_color_manual(values = cbpal[c(2,3)],guide = guide_legend(override.aes = list(alpha=1))) +
  scale_y_continuous(breaks = seq(0,250,by=50)) +
  labs(y = "monthly precipitation (mm)", x = "date") +
  theme_minimal()

prets_plot2

ggsave(here("figures","mordor-ab-monthly-precip.png"),device = "png", width = 8, height = 4)

```

# Save a precipitation dataset

```{r save precipitation data}
d_pre5 <- d_pre4 %>%
  ungroup() %>%
  arrange(clusterid,pdate) %>%
  dplyr::select(clusterid, pdate, pre, season)

write_csv(d_pre5, path = here("data","mordor-ab-monthly-precip.csv"))

```


# Session Info
```{r session info}
sessionInfo()
```


